---
output: html_document
editor_options: 
  chunk_output_type: console
---
0---
title: Report on data in AusTraits
output:
  html_document:
    df_print: kable
    highlight: tango
    keep_md: no
    smart: no
    theme: yeti
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options:
  chunk_output_type: console
---

Load data
```{r, warning=FALSE, message=FALSE}
# Install dependencies as needed
# devtools::install_deps()
# Load packages and functions for this analysis

rm(list=ls())
# devtools::load_all(quiet = TRUE)
source("R/calc_c13_a.R")
source("R/error_checking_function.R")
source("R/plot_trait_by_dataset.R")
source("R/figure_opt_plot.R")
source("R/loading data.R")
source("R/plot_whittaker_biomes.R")
source("R/plot_climate.R")

library(xtable)
library(lubridate)
library(R.utils)
library(tidyverse)
library(corrplot)
library(ggpointdensity)
library(tagger)
library(relaimpo)

# load trait data
austraits <- readRDS("data/austraits-4.0.0.rds")
```

```{r}
location_of_sites <-
  austraits$locations %>%
  filter(location_property %in% c("longitude (deg)", "latitude (deg)")) %>%
  spread(location_property, value) %>%
  rowid_to_column("ID") %>%
  rename(latitude = `latitude (deg)`, longitude = `longitude (deg)`) %>%
  mutate_at(c("latitude", "longitude"), ~suppressWarnings(as.numeric(.x))) %>%
  drop_na(longitude, latitude)

```


```{r}
sum_no_NA <- function(x) if(all(is.na(x))) NA_integer_ else sum(x, na.rm=TRUE)
```


```{r}
#where are sites located?
location_of_sites <-
  austraits$locations %>%
  filter(location_property %in% c("longitude (deg)", "latitude (deg)")) %>%
  spread(location_property, value) %>%
  rowid_to_column("ID") %>%
  rename(latitude = `latitude (deg)`, longitude = `longitude (deg)`) %>%
  mutate_at(c("latitude", "longitude"), ~suppressWarnings(as.numeric(.x))) %>%
  drop_na(longitude, latitude)


# load map of australia (required for climate data analysis)
au_map <- raster::raster("data/australia.tif")
```


```{r}
traits <- austraits$traits %>%
  left_join(location_of_sites) %>%
  filter(trait_name %in% c("leaf_area","leaf_dry_mass","leaf_mass_per_area","leaf_N_per_dry_mass","leaf_N_per_area","wood_density","plant_height", "seed_dry_mass", "leaf_delta13C","huber_value")) %>%
  mutate(value = as.numeric(value))
```

Create a new trait dataframe to only include observations with associated latitude and longitude
```{r}
traits_georef <- traits %>%
  drop_na(latitude,longitude)
```

```{r}
traits_georef %>%
  filter(basis_of_record %in% c("field", "field, field_experiment")) %>%
  filter(!(dataset_id == "Cheesman_2020" & (method_id %in% c("03","04") | treatment_id %in% c("03")))) -> field_traits_georef
```

We also remove all trait values which are records across multiple sites, are from the literature, expert values, or are from unknown sources

```{r}
field_traits_georef %>%
  filter(basis_of_value %in% c("measurement") | trait_name == "plant_height") -> field_traits_georef
```


For species with more than one observation, we can check to see if their values are relatively consistent by inspecting max:min ratio, sd, CV etc. We need to do this seperately for measurements involving leaves and for the rest of the traits because of the fact that hte leaf vs. leaflet perspective will lead to large differences in measurements. So we need to first divide the leaf area dataset up into leaf area measurements on simple leaves, compound leaves measured at the leaflet scale, and leaf area measured at the leaf scale. 


```{r}
read_csv("data/datasets_with_leaf_area_or_dry_mass.csv") %>%
  dplyr::select(dataset_id, compound_species, leaf_unit_measured) %>%
  right_join(field_traits_georef) %>%
  filter(trait_name %in% c("leaf_area","leaf_dry_mass")) -> field_traits_georef_leaf_area_leaf_dry_mass

field_traits_georef_leaf_area_leaf_dry_mass %>%
  filter(compound_species == "no_compound") -> no_compound

field_traits_georef_leaf_area_leaf_dry_mass %>%
  filter(compound_species == "compound_present" & leaf_unit_measured == "leaflet") -> compound_leaflet

field_traits_georef_leaf_area_leaf_dry_mass%>%
  filter(compound_species == "compound_present" & leaf_unit_measured %in% c(NA,"leaf"))-> compound_leaf

austraits$traits %>%
  filter(trait_name == "leaf_compoundness") %>%
  group_by(taxon_name) %>%
  mutate(num_types = n_distinct(value)) %>%
  ungroup() %>%
  mutate(compoundness = if_else(num_types > 1 | value =="compound" | value == "compound simple", "compound", "simple")) %>% 
  dplyr::select(-value) %>%
  distinct(taxon_name, .keep_all = T) %>%
  dplyr::select(compoundness, taxon_name) %>%
  right_join(compound_leaf) -> compound_leaf_remove_compound

compound_leaf_remove_compound %>%
  filter(is.na(compoundness)) %>% 
  pull(taxon_name) %>%
  unique()-> unknown_species

read_csv("data/leaf_area_taxa_taxonomic_master.csv") %>%
  drop_na(leaf_description) %>%
  filter(taxon_name %in% unknown_species) %>%
  dplyr::select(taxon_name, leaf_description) %>%
  rename(compoundness = leaf_description) %>%
  right_join(compound_leaf_remove_compound, by = "taxon_name")%>%
  mutate(compoundness = coalesce(compoundness.x, compoundness.y)) %>%
  filter(!compoundness %in% c("compound")) %>%
  drop_na(compoundness) %>%
  distinct(taxon_name) -> species_to_keep
  
compound_leaf %>%
  filter(taxon_name %in% species_to_keep$taxon_name) -> compound_leaf

field_traits_georef_leaf_area_leaf_dry_mass <- bind_rows(no_compound, compound_leaflet, compound_leaf) %>%
  dplyr::select(-leaf_unit_measured, -compound_species)
```

```{r}
field_traits_georef %>%
  filter(!trait_name %in% c("leaf_area","leaf_dry_mass")) %>%
  bind_rows(field_traits_georef_leaf_area_leaf_dry_mass) -> field_traits_georef
```


```{r}
austraits$traits %>%
  filter(trait_name %in% c("plant_growth_form"))  %>% 
  filter(dataset_id == "Wenk_2022") %>%
  filter(value %in% c("mallee", 
                      "mallee shrub", 
                      "mallee shrub tree", 
                      "mallee tree", 
                      "shrub", 
                      "shrub subshrub",
                      "shrub tree",
                      "subshrub",
                      "tree")) %>%
  pull(taxon_name) %>%
  unique() -> woody_taxa

austraits$traits %>%
  filter(trait_name %in% c("plant_growth_form"))  %>% 
  filter(dataset_id == "Wenk_2022") %>%
  filter(!value %in% c("mallee", 
                      "mallee shrub", 
                      "mallee shrub tree", 
                      "mallee tree", 
                      "shrub", 
                      "shrub subshrub",
                      "shrub tree",
                      "subshrub",
                      "tree")) %>%
  pull(taxon_name) %>%
  unique() -> non_woody_taxa



austraits$traits %>%
  filter(trait_name %in% c("plant_growth_form"))  %>% 
  filter(dataset_id == "Wenk_2022") %>%
  filter(value %in% c("climber shrub", 
                      "climber_woody", 
                      "climber_woody shrub", 
                      "climber_woody shrub subshrub", 
                      "climber_woody shrub tree", 
                      "climber_woody subshrub",
                      "climber_woody tree")) %>%
  pull(taxon_name) %>%
  unique() -> woody_climbers


austraits$traits %>%
  filter(trait_name %in% c("plant_growth_form"))  %>% 
  filter(dataset_id == "Wenk_2022") %>%
  filter(!value %in% c("climber shrub", 
                      "climber_woody", 
                      "climber_woody shrub", 
                      "climber_woody shrub subshrub", 
                      "climber_woody shrub tree", 
                      "climber_woody subshrub",
                      "climber_woody tree")) %>% 
  pull(taxon_name) %>%
  unique() -> not_woody_climbers




field_traits_georef %>%
  mutate(growth_form = case_when(taxon_name %in% woody_taxa ~ "Woody",
                                 taxon_name %in% non_woody_taxa ~ "Non-woody")) -> woody_field_traits_georef

woody_field_traits_georef %>%
  mutate(woody_climber = case_when(taxon_name %in% woody_climbers ~ "Yes",
                                 taxon_name %in% not_woody_climbers ~ "No")) -> woody_field_traits_georef

woody_field_traits_georef %>% 
  filter(is.na(growth_form)) %>% 
  distinct(taxon_name) -> taxa_without_growth_form

write.csv(taxa_without_growth_form, "data/taxa_without_growth_form.csv")
```

```{r}
woody_field_traits_georef %>%
  filter(trait_name %in% c("leaf_N_per_dry_mass", "leaf_mass_per_area", "leaf_N_per_area")) %>%
  mutate(value = as.numeric(value)) %>%
  group_by(dataset_id, taxon_name, observation_id, trait_name, ID, latitude, longitude, growth_form, location_name, woody_climber) %>% 
  summarise(value = mean(value, na.rm = T)) %>%
  ungroup() %>%
  pivot_wider(names_from = "trait_name", values_from = "value") %>%
  mutate(leaf_N_per_dry_mass = leaf_N_per_dry_mass/1000) -> pre_calculated_leaf_N_data

pre_calculated_leaf_N_data %>% 
  filter(is.na(leaf_N_per_area) &
           !is.na(leaf_N_per_dry_mass) &
           !is.na(leaf_mass_per_area)) %>%
  nrow()

pre_calculated_leaf_N_data %>%
  mutate(leaf_N_per_area= if_else(!is.na(leaf_N_per_area), 
                                  leaf_N_per_area, 
                                  leaf_N_per_dry_mass * leaf_mass_per_area)) %>%
  dplyr::select(-leaf_N_per_dry_mass, -leaf_mass_per_area) %>%
  drop_na(leaf_N_per_area) %>%
  pivot_longer(cols = leaf_N_per_area, names_to = "trait_name") %>%
  mutate(trait_name = "leaf_N_per_area_calc") -> calculated_leaf_N_data

woody_field_traits_georef %>%
  bind_rows(calculated_leaf_N_data) -> woody_field_traits_georef
```

Error checking for leaf mass per area

```{r}
error_check(woody_field_traits_georef, "leaf_mass_per_area") -> error_check_lma
```

Error checking for huber value

```{r}
error_check(woody_field_traits_georef, "huber_value") -> error_check_huber
```

Error checking for leaf N per area

```{r}
error_check(woody_field_traits_georef, "leaf_N_per_area") -> error_check_leaf_N
```

Error checking for wood density - no problematic data points detected

```{r}
error_check(woody_field_traits_georef, "wood_density") -> error_check_wood_density
```

Error checking for plant height - no problematic data points detected

```{r}
error_check(woody_field_traits_georef, "plant_height") -> error_check_plant_height
```


Error checking for leaf_delta13C - no problematic points detected

```{r}
error_check(woody_field_traits_georef, "leaf_delta13C") -> error_check_13C
```

Error checking for seed mass

```{r}
error_check(woody_field_traits_georef, "seed_dry_mass") -> error_check_seed_mass
```

Error checking for leaf area

```{r}
error_check(woody_field_traits_georef, "leaf_area") -> error_check_leaf_area
```

To see if there any instances where data is shared between studies, we inspect occurrences of identical trait values for each trait~species group. Although there are a number of cases where this occurs, most seem to be coincidental.

```{r}
woody_field_traits_georef %>%
  group_by(value, taxon_name, trait_name) %>% 
  filter(n()>1) %>% 
  filter(n_distinct(dataset_id)>1) -> identical_values
```


```{r}
erroneous_values <- read_csv("data/erroneous_values.csv")

erroneous_list <- vector("list", nrow(erroneous_values))

for(i in 1:nrow(erroneous_values)){
  erroneous_data <- erroneous_values[i,] %>% dplyr::select_if(~ !any(is.na(.)))
  erroneous_data <- woody_field_traits_georef %>% left_join(erroneous_data) %>% drop_na(erroneous)
  erroneous_list[[i]] <- erroneous_data
}

woody_field_traits_georef <- left_join(woody_field_traits_georef,bind_rows(erroneous_list))

woody_field_traits_georef %>%
  filter(is.na(erroneous)) -> woody_field_traits_georef
```

```{r}
woody_field_traits_georef %>%
  mutate(value = case_when(trait_name == "plant_height" & dataset_id == "Dwyer_2017" ~ value/10,
                           TRUE ~ as.numeric(value))) -> woody_field_traits_georef
```

```{r}
woody_field_traits_georef %>%
  filter(trait_name == "leaf_delta13C") %>%
  mutate(trait_name = "leaf_capital_delta13C") %>%
  mutate(value = (-8/1000 - value/1000)/(1 + value/1000)*1000) -> leaf_Delta13C_calculated
```

```{r}
woody_field_traits_georef %>%
  bind_rows(leaf_Delta13C_calculated) -> woody_field_traits_georef
```


```{r}
#define the core traits to focus on
core_traits <- c("huber_value","leaf_capital_delta13C","leaf_N_per_area_calc","leaf_mass_per_area","wood_density","plant_height","leaf_area","seed_dry_mass")
```

```{r}
woody_field_traits_georef %>%
  filter(trait_name %in% core_traits) %>%
    filter(!trait_name %in% c("huber_value","wood_density")) %>%
  plot_site_map_by_trait() -> overall_distribution

woody_field_traits_georef %>%
  filter(growth_form == "Woody") %>%
  filter(trait_name %in% core_traits) %>%
  plot_site_map_by_trait() -> woody_distribution

woody_field_traits_georef %>%
  filter(growth_form != "Woody") %>%
  filter(trait_name %in% core_traits) %>%
  filter(!trait_name %in% c("huber_value","wood_density")) %>%
  plot_site_map_by_trait() -> non_woody_distribution

png("output/manuscript/overall_distribution.png", height = 1333, width = 2000, res = 300)
overall_distribution +
  labs(title = "All trait data")
dev.off()

png("output/manuscript/woody_distribution.png", height = 2000, width = 2000, res = 300)
woody_distribution +
  labs(title = "Woody trait data")
dev.off()

png("output/manuscript/non_woody_distribution.png", height = 1333, width = 2000, res = 300)
non_woody_distribution +
  labs(title = "Non-woody trait data")
dev.off()

```

#Run from above

```{r}
source("R/loading data_test.R")
climate_data <- load_climate_data(woody_field_traits_georef)
```

#Read in climate data

```{r}
woody_field_traits_georef %>%
  bind_rows(leaf_Delta13C_calculated) %>%
  left_join(climate_data) %>% 
  group_by(cell, taxon_name, trait_name, growth_form, woody_climber, elev, prec, srad, tavg , tmax, prec.cv, temp.cv, vpd,prec.wq, prec.dq,aridityIndexThornthwaite, climaticMoistureIndex, growingDegDays5) %>%
  summarise(value=mean(value, na.rm=T), latitude = mean(latitude, na.rm = T), longitude = mean(latitude, na.rm = T)) %>%
  ungroup() -> woody_field_traits_georef_mean_climate

woody_field_traits_georef_mean_climate %>% View()
```

We will now inspect distribution of data in each study relative to total distribution to see if there is any erroneous data
```{r eval=FALSE, include=FALSE}
traits_to_inspect <- woody_field_traits_georef_tree_form %>%
   distinct(trait_name)

trait_by_dataset_core_traits<-function(trait){

woody_field_traits_georef_tree_form %>%
  filter(trait_name == trait) %>%
  distinct(dataset_id) -> dataset_index

purrr::map(dataset_index$dataset_id, plot_trait_by_dataset, trait)
}

purrr::map(traits_to_inspect$trait_name, trait_by_dataset_core_traits)
```

#Prep data for analysis
First, let's calculate the correlations between the traits and the environmental variables. This is will be the key statistic which will be used for analysis.For starters, which traits and environmental variables are not normally distributed? We need to turn the data into a wide format.

```{r}
woody_field_traits_georef_mean_climate %>%
  filter(trait_name %in% core_traits) %>%
  dplyr::select(trait_name, value, Temp = tavg, SWdown = srad, Prec = prec, Max_temp = tmax, elev = elev,VPD = vpd, prec_cv = prec.cv, temp_cv = temp.cv, prec_wq = prec.wq, prec_dq = prec.dq, AI_thorn = aridityIndexThornthwaite, MI = climaticMoistureIndex, GDD_5 =  growingDegDays5, taxon_name, growth_form, woody_climber, cell) %>%
  pivot_wider(values_from = value, names_from = trait_name) -> woody_field_traits_georef_mean_climate_wide


woody_field_traits_georef_mean_climate_wide %>%
  mutate(log_leaf_capital_delta13C = log(leaf_capital_delta13C),
         log_leaf_area = log(leaf_area),
         log_seed_mass = log(seed_dry_mass),
         log_LMA = log(leaf_mass_per_area),
         log_wood_dens = log(wood_density),
         log_leaf_N = log(leaf_N_per_area_calc),
         log_height = log(plant_height),
         log_huber = log(huber_value),
         log_VPD = log(VPD),
         log_Prec = log(Prec),
         log_prec_wq = log(prec_wq),
         sqrt_prec_dq = sqrt(prec_dq),
          log_Prec_CV = log(prec_cv),
         sqrt_elev = sqrt(elev)) -> woody_field_traits_georef_mean_climate_wide

#environmental variables
hist(woody_field_traits_georef_mean_climate_wide$Temp)
hist(woody_field_traits_georef_mean_climate_wide$GDD_5)
hist(log(woody_field_traits_georef_mean_climate_wide$VPD))
hist(woody_field_traits_georef_mean_climate_wide$SWdown)
hist(log(woody_field_traits_georef_mean_climate_wide$Prec))
hist(sqrt(woody_field_traits_georef_mean_climate_wide$prec_dq))
hist(log(woody_field_traits_georef_mean_climate_wide$prec_wq))

hist(sqrt(woody_field_traits_georef_mean_climate_wide$elev))
hist(woody_field_traits_georef_mean_climate_wide$Max_temp)
hist(woody_field_traits_georef_mean_climate_wide$MI)
hist(woody_field_traits_georef_mean_climate_wide$AI_thorn)
hist(log(woody_field_traits_georef_mean_climate_wide$prec_cv))
hist((woody_field_traits_georef_mean_climate_wide$temp_cv))


#traits
hist(log(woody_field_traits_georef_mean_climate_wide$leaf_area))
hist(log(woody_field_traits_georef_mean_climate_wide$leaf_mass_per_area))
hist((woody_field_traits_georef_mean_climate_wide$wood_density))
hist(log(woody_field_traits_georef_mean_climate_wide$leaf_N_per_area_calc))
hist(log(woody_field_traits_georef_mean_climate_wide$plant_height))
hist(log(woody_field_traits_georef_mean_climate_wide$huber_value))
hist(log(woody_field_traits_georef_mean_climate_wide$seed_dry_mass))
hist(woody_field_traits_georef_mean_climate_wide$leaf_capital_delta13C)

#all traits bar wood density will be log transformed.
```

```{r}
bioclim <- raster::stack(list.files("data/climate_data/wc2.1_30s_bio", full.names = T)[c(1,4,14)])

bioclim_au <-bioclim %>%
  raster::projectRaster(au_map) %>%
  raster::mask(au_map)

global_values_temp_seas <- tibble(temp_seas = raster::values(bioclim[[3]]))
au_values_temp_seas <- tibble(temp_seas = raster::values(bioclim_au[[3]]))

au_values_temp_seas %>%
  summarise(min_temp_seas = min(temp_seas, na.rm = T),
            max_temp_seas = max(temp_seas, na.rm = T)) -> au_values_temp_seas_summarised

global_values_temp_seas %>% drop_na() %>%
  pull(temp_seas) %>% ecdf() -> global_values_temp_seas_ecdf

global_values_temp_seas_ecdf(au_values_temp_seas_summarised$min_temp_seas)
global_values_temp_seas_ecdf(au_values_temp_seas_summarised$max_temp_seas)

global_values_temp_seas %>%
  sample_n(1000000) %>%
  drop_na() %>%
  ggplot() +
  ylab("Quantile") +
  xlab("Temperature Seasonality") +
  theme_classic() +
  geom_polygon(data = tibble(y = c(0,1,1,0),
                             x = c(au_values_temp_seas_summarised$min_temp_seas,
                                         au_values_temp_seas_summarised$min_temp_seas,
                                         au_values_temp_seas_summarised$max_temp_seas,
                                         au_values_temp_seas_summarised$max_temp_seas)),
                             aes(x = x, y = y), alpha = 0.3, fill = "red") +
  stat_ecdf(aes(temp_seas), linewidth = 2) + 
  theme(text=element_text(size=12)) -> ecdf_temp_seas

rm(global_values_temp_seas)

global_values_temp <- tibble(temp = raster::values(bioclim[[1]]))
au_values_temp <- tibble(temp = raster::values(bioclim_au[[1]]))

au_values_temp %>%
  summarise(min_temp = min(temp, na.rm = T),
            max_temp = max(temp, na.rm = T)) -> au_values_temp_summarised

global_values_temp %>% drop_na() %>%
  pull(temp) %>% ecdf() -> global_values_temp_ecdf

global_values_temp_ecdf(au_values_temp_summarised$min_temp)
global_values_temp_ecdf(au_values_temp_summarised$max_temp)


global_values_temp %>%
  drop_na() %>%
  sample_n(1000000) %>%
  ggplot() +
  ylab("Quantile") +
  xlab(expression(Mean~annual~temperature~(degree * C))) +
  theme_classic() +
  geom_polygon(data = tibble(y = c(0,1,1,0),
                             x = c(au_values_temp_summarised$min_temp,
                                         au_values_temp_summarised$min_temp,
                                         au_values_temp_summarised$max_temp,
                                         au_values_temp_summarised$max_temp)),
                             aes(x = x, y = y), alpha = 0.3, fill = "red") +
  stat_ecdf(aes(temp), linewidth = 2) + 
  theme(text=element_text(size=12)) -> ecdf_temp

rm(global_values_temp)

global_values_prec <- tibble(prec = raster::values(bioclim[[2]]))
au_values_prec <- tibble(prec = raster::values(bioclim_au[[2]]))

au_values_prec %>%
  summarise(min_prec = min(prec, na.rm = T),
            max_prec = max(prec, na.rm = T)) -> au_values_prec_summarised

global_values_prec %>% drop_na() %>%
  pull(prec) %>% ecdf() -> global_values_prec_ecdf

global_values_prec_ecdf(au_values_prec$min_prec)
global_values_prec_ecdf(au_values_prec$max_prec)


global_values_prec %>%
  drop_na() %>%
  sample_n(1000000) %>%
  ggplot() +
  ylab("Quantile") +
  xlab("Mean annual precipitation (mm)") +
  theme_classic() +
  geom_polygon(data = tibble(y = c(0,1,1,0),
                             x = c(au_values_prec_summarised$min_prec,
                                         au_values_prec_summarised$min_prec,
                                         au_values_prec_summarised$max_prec,
                                         au_values_prec_summarised$max_prec)),
                             aes(x = x, y = y), alpha = 0.3, fill = "red") +
  stat_ecdf(aes(prec), linewidth = 2) + 
  theme(text=element_text(size=12)) -> ecdf_prec

rm(global_values_prec)

dev.off()
source("R/plot_whittaker_biomes.R")
whittaker_plots <- austraits_climate_space(core_traits)
second_column <-cowplot::plot_grid(ecdf_prec, ecdf_temp, ecdf_temp_seas, labels = c("b","c","d"), ncol = 1)
combined_plots <- cowplot::plot_grid(whittaker_plots, second_column, labels = "a","", nrow = 1, ncol = 2)

png("output/manuscript/ecdf_whittaker.png", height = 1800, width = 2800, res = 300)
combined_plots
dev.off()
cor(au_values_prec, au_values_temp, use = "pairwise.complete.obs")
```


First, just check relationships between environmental variables
```{r}
png("output/manuscript/supps_ellipse_environmental.png", height=1300, width=2000, res=300)

library(corrplot)

woody_field_traits_georef_mean_climate_wide %>%
  drop_na(growth_form) %>%
  distinct(cells, .keep_all = T) %>%
  dplyr::select(-c(taxon_name,cells, growth_form, woody_climber)) -> for_corrplot
trait_env_correlation<- cor(for_corrplot, use="pairwise.complete.obs")

corrplot_example <- trait_env_correlation[c("Temp","GDD_5","log_VPD","SWdown","log_Prec","sqrt_elev","Max_temp", "MI","AI_thorn","log_Prec_CV","temp_cv","log_prec_wq","sqrt_prec_dq"),c("Temp","GDD_5","log_VPD","SWdown","log_Prec","sqrt_elev","Max_temp", "MI","AI_thorn","log_Prec_CV","temp_cv","log_prec_wq","sqrt_prec_dq")]

colnames(corrplot_example) <- c("MAT","GDD_5","log(VPD)","SWDown","log(MAP)","sqrt(Elevation)","max(Temp)", "climatic moisture index", "aridity index Thornthwaite","log_Prec_CV","temp_cv","Prec_WQ", "Prec_DQ")
rownames(corrplot_example) <- c("MAT","GDD_5","log(VPD)","SWDown","log(MAP)","sqrt(Elevation)","max(Temp)", "climatic moisture index", "aridity index Thornthwaite","log_Prec_CV","temp_cv","Prec_WQ", "Prec_DQ")

corrplot::corrplot(corrplot_example, method="ellipse", type = "upper")

dev.off()

trait_specific_correlations <- function(trait){
woody_field_traits_georef_mean_climate_wide %>%
  drop_na(trait) %>%
  distinct(cells, .keep_all = T) %>%
  dplyr::select(-c(taxon_name,cells, growth_form, woody_climber)) -> for_corrplot
trait_env_correlation<- cor(for_corrplot, use="pairwise.complete.obs")

corrplot_example <- trait_env_correlation[c("Temp","GDD_5","log_VPD","SWdown","log_Prec","sqrt_elev","Max_temp", "climaticMoistureIndex","aridityIndexThornthwaite","log_MI","log_Prec_CV","temp_cv"),c("Temp","GDD_5","log_VPD","SWdown","log_Prec","sqrt_elev","Max_temp", "climaticMoistureIndex","aridityIndexThornthwaite","log_MI","log_Prec_CV","temp_cv")]

colnames(corrplot_example) <- c("MAT","GDD_5","log(VPD)","SWDown","log(MAP)","sqrt(Elevation)","max(Temp)", "climatic moisture index", "aridity index Thornthwaite", "log(MI)","log_Prec_CV","temp_cv")
rownames(corrplot_example) <- c("MAT","GDD_5","log(VPD)","SWDown","log(MAP)","sqrt(Elevation)","max(Temp)", "climatic moisture index", "aridity index Thornthwaite", "log(MI)","log_Prec_CV","temp_cv")

p <- corrplot::corrplot(corrplot_example, method="ellipse", type = "upper")
p
}

 map(woody_field_traits_georef_mean_climate %>% filter(!trait_name %in% c("leaf_delta13C", "leaf_N_per_area", "leaf_N_per_dry_mass", "leaf_dry_mass")) %>%  pull(trait_name) %>% unique(), trait_specific_correlations)
```

Basic patterns with environment

Supp Figure. 1

```{r}
fit_models <- function(...){
  data <- tibble(...) %>%
    unnest(data)
  fit_lm <- lm(trait_value~env_value, data)
  
  data$env_value2 <- data$env_value^2
  fit_quad <- lm(trait_value ~ env_value + env_value2, data)
  
  return(tibble(p_lm = summary(fit_lm)$coef[2,4],
         fit_lm = summary(fit_lm)$r.squared, 
         fit_quad = summary(fit_quad)$r.squared))
}


woody_field_traits_georef_mean_climate_wide %>%
  drop_na(growth_form) %>%
  dplyr::select(log_leaf_capital_delta13C, log_height, log_leaf_N, log_LMA, log_seed_mass, log_leaf_area, log_Prec, Temp, GDD_5, SWdown, Max_temp, elev, log_VPD, MI, AI_thorn,log_Prec_CV,temp_cv, log_prec_wq, sqrt_prec_dq) %>%
  pivot_longer(cols = c(log_leaf_capital_delta13C, 
                        log_leaf_area,
                        log_seed_mass,
                        log_LMA,
                        log_leaf_N,
                        log_height),
                        names_to = "trait_name", values_to = "trait_value") %>%
  pivot_longer(cols = c(log_Prec, Temp, GDD_5, SWdown, Max_temp, elev, log_VPD, MI, AI_thorn,log_Prec_CV,temp_cv, log_prec_wq, sqrt_prec_dq), names_to = "env_name", values_to = "env_value") %>% 
  group_by(trait_name, env_name) %>%
  nest() %>%
  ungroup %>%
  mutate(r2 = pmap(., fit_models)) %>%
  unnest(r2) -> overall_relationships

overall_relationships %>%
  dplyr::select(fit_lm, trait_name, env_name) %>%
  pivot_wider(names_from = "env_name", values_from ="fit_lm") -> fit_lm_overall

print(xtable(fit_lm_overall %>% pivot_longer(cols = -(trait_name)) %>% pivot_wider(names_from = "trait_name", values_from = "value")), include.rownames=FALSE, file = "output/manuscript/fit_lm_overall.tex")

overall_relationships %>%
  dplyr::select(fit_quad, trait_name, env_name) %>%
  pivot_wider(names_from = "env_name", values_from ="fit_quad") -> fit_quad_overall

print(xtable(fit_lm_overall %>% pivot_longer(cols = -(trait_name)) %>% pivot_wider(names_from = "trait_name", values_from = "value")), include.rownames=FALSE, file = "output/manuscript/fit_quad_overall.tex")

  woody_field_traits_georef_mean_climate_wide %>% 
  filter(growth_form == "Woody" & woody_climber == "No") %>%
  dplyr::select(log_huber, log_wood_dens, log_leaf_capital_delta13C, log_height, log_leaf_N, log_LMA, log_seed_mass, log_leaf_area, log_Prec, Temp, GDD_5, SWdown, Max_temp, elev, log_VPD, MI, AI_thorn,log_Prec_CV,temp_cv, log_prec_wq, sqrt_prec_dq) %>%
  pivot_longer(cols = c(log_leaf_capital_delta13C, 
                        log_leaf_area,
                        log_seed_mass,
                        log_LMA,
                        log_leaf_N,
                        log_height,
                        log_huber,
                        log_wood_dens),
                        names_to = "trait_name", values_to = "trait_value") %>%
  pivot_longer(cols = c(log_Prec, Temp, GDD_5, SWdown, Max_temp, elev, log_VPD, MI, AI_thorn,log_Prec_CV,temp_cv, log_prec_wq, sqrt_prec_dq), names_to = "env_name", values_to = "env_value") %>% 
  group_by(trait_name, env_name) %>%
  nest() %>%
  ungroup %>%
  mutate(r2 = pmap(., fit_models)) %>%
  unnest(r2) -> woody_relationships

woody_relationships %>%
  dplyr::select(fit_lm, trait_name, env_name) %>%
  pivot_wider(names_from = "env_name", values_from ="fit_lm") -> fit_lm_woody

print(xtable(fit_lm_woody %>% pivot_longer(cols = -(trait_name)) %>% pivot_wider(names_from = "trait_name", values_from = "value")), include.rownames=FALSE, file = "output/manuscript/fit_lm_woody.tex")

woody_relationships %>%
  dplyr::select(fit_quad, trait_name, env_name) %>%
  pivot_wider(names_from = "env_name", values_from ="fit_quad") -> fit_quad_woody

print(xtable(fit_quad_woody %>% pivot_longer(cols = -(trait_name)) %>% pivot_wider(names_from = "trait_name", values_from = "value")), include.rownames=FALSE, file = "output/manuscript/fit_quad_woody.tex")

woody_field_traits_georef_mean_climate_wide %>%
  filter(growth_form == "Non-woody") %>%
  dplyr::select(log_leaf_capital_delta13C, log_height, log_leaf_N, log_LMA, log_seed_mass, log_leaf_area, log_Prec, Temp, GDD_5, SWdown, Max_temp, elev, log_VPD, MI, AI_thorn,log_Prec_CV,temp_cv, log_prec_wq, sqrt_prec_dq) %>%
  pivot_longer(cols = c(log_leaf_capital_delta13C, 
                        log_leaf_area,
                        log_seed_mass,
                        log_LMA,
                        log_leaf_N,
                        log_height),
                        names_to = "trait_name", values_to = "trait_value") %>%
  pivot_longer(cols = c(log_Prec, Temp, GDD_5, SWdown, Max_temp, elev, log_VPD, MI, AI_thorn,log_Prec_CV,temp_cv, log_prec_wq, sqrt_prec_dq), names_to = "env_name", values_to = "env_value") %>% 
  group_by(trait_name, env_name) %>%
  nest() %>%
  ungroup %>%
  mutate(r2 = pmap(., fit_models)) %>%
  unnest(r2)  -> non_woody_relationships
  
non_woody_relationships %>%
  dplyr::select(fit_lm, trait_name, env_name) %>%
  pivot_wider(names_from = "env_name", values_from ="fit_lm") -> fit_lm_non_woody

print(xtable(fit_lm_non_woody %>% pivot_longer(cols = -(trait_name)) %>% pivot_wider(names_from = "trait_name", values_from = "value")), include.rownames=FALSE, file = "output/manuscript/fit_lm_non_woody.tex")

non_woody_relationships %>%
  dplyr::select(fit_quad, trait_name, env_name) %>%
  pivot_wider(names_from = "env_name", values_from ="fit_quad") -> fit_quad_non_woody

print(xtable(fit_quad_non_woody %>% pivot_longer(cols = -(trait_name)) %>% pivot_wider(names_from = "trait_name", values_from = "value")), include.rownames=FALSE, file = "output/manuscript/fit_quad_non_woody.tex")
```


```{r}
woody_field_traits_georef_mean_climate %>%
    filter(growth_form == "Woody" & woody_climber == "No") %>%
    dplyr::select(trait_name, value, prec, growth_form)%>%
    pivot_longer(-c("trait_name", "value","growth_form"), names_to = "env", values_to = "env_value")%>%
    # left_join(sub_labels) %>%
    filter(trait_name %in% c("leaf_capital_delta13C","leaf_mass_per_area","leaf_N_per_area_calc","wood_density", "seed_dry_mass","huber_value")) %>%
    group_by(trait_name) %>%
    mutate(cor = round(cor(log(value), log(env_value), use = "pairwise.complete.obs"), digits = 2)) %>%
    ungroup() %>%
    drop_na(value, env_value) %>%
    group_by(trait_name) %>%
    mutate(n = n()) %>%
    ungroup() -> data_for_plot
  
  data_for_plot$trait_name <- factor(data_for_plot$trait_name, labels = c("H[v]", "Delta^13~C~(`\u2030`)", "LMA~(g~m^{-2})", "N[area]~(g~m^{-2})", "SM~(mg)" , "Wood~density~(mg~mm^{-3})"))

data_for_plot %>%
  ggplot(aes(env_value, value,group = trait_name)) +
  geom_point(alpha = 0.5, col = "#db6d00") + 
  scale_y_log10() +
  scale_x_log10(limits = c(min(woody_field_traits_georef_mean_climate$prec, na.rm = T), max(woody_field_traits_georef_mean_climate$prec, na.rm = T))) +
  theme_classic() +
  facet_wrap(~trait_name, scales = "free_y", 
                strip.position = "left",
               labeller = label_parsed) +
     theme(strip.background = element_blank(),
           strip.placement = "outside") +
  xlab("MAP (mm)") +
  ylab("") +
  theme(text=element_text(size=20), legend.position = "none") +
    # geom_text(mapping = aes(x = 0, y = Inf, label = label, group=env),
    #         hjust = "inward", vjust = "inward",
    #         inherit.aes = FALSE, check_overlap = TRUE, size = 6) +
  geom_text(mapping = aes(x = 0, y = 0, label = paste("r = ",cor, " n = ", n)), inherit.aes = FALSE, size = 4.5, vjust = -1, hjust = -0.125, check_overlap = T) +
  geom_smooth(method = "lm", col = "black", alpha = 0.8) +
  tag_facets() ->p
  
    
  png("output/manuscript/woody_relationships_prec_specific.png", height = 2000, width = 4000, res = 250)
  p
  dev.off()

woody_field_traits_georef_mean_climate %>%
    dplyr::select(trait_name, value, prec, growth_form)%>%
  filter(growth_form %in% c("Woody", "Non-woody")) %>%
    pivot_longer(-c("trait_name", "value","growth_form"), names_to = "env", values_to = "env_value")%>%
    # left_join(sub_labels) %>%
    filter(trait_name %in% c("plant_height","leaf_area")) %>%
    group_by(trait_name) %>%
    mutate(cor = round(cor(log(value), log(env_value), use = "pairwise.complete.obs"), digits = 2)) %>%
    ungroup() %>%
    drop_na(value, env_value) %>%
    group_by(trait_name) %>%
    mutate(n = n()) %>%
    ungroup() -> data_for_plot
  
  data_for_plot$trait_name <- factor(data_for_plot$trait_name, labels = c("LA~(mm^2)", "H~(m)"))

data_for_plot %>%
  ggplot(aes(env_value, value,group = trait_name)) +
  geom_point(aes(group = growth_form, colour = growth_form), alpha = 0.5) + 
  scale_y_log10() +
  scale_x_log10(limits = c(min(woody_field_traits_georef_mean_climate$prec, na.rm = T), max(woody_field_traits_georef_mean_climate$prec, na.rm = T))) +
  theme_classic() +
  facet_wrap(~trait_name, scales = "free_y", 
                strip.position = "left",
               labeller = label_parsed) +
     theme(strip.background = element_blank(),
           strip.placement = "outside") +
  xlab("MAP (mm)") +
  ylab("") +
  scale_color_manual(values = c("#009292","#db6d00")) +
  theme(text=element_text(size=20), legend.position = "none") +
    # geom_text(mapping = aes(x = 0, y = Inf, label = label, group=env),
    #         hjust = "inward", vjust = "inward",
    #         inherit.aes = FALSE, check_overlap = TRUE, size = 6) +
  geom_text(mapping = aes(x = 0, y = 0, label = paste("r = ",cor, " n = ", n)), inherit.aes = FALSE, size = 4.5, vjust = -1, hjust = -0.125, check_overlap = T) +
  geom_smooth(method = "lm", col = "black", alpha = 0.8) +
  tag_facets() ->p
  
    
  png("output/manuscript/overall_relationships_prec_specific.png", height = 1200, width = 3333, res = 250)
  p
  dev.off()  
  
```

```{r}
woody_field_traits_georef_mean_climate %>%
  drop_na(growth_form) %>%
    dplyr::select(trait_name, value, prec, growth_form)%>%
    pivot_longer(-c("trait_name", "value","growth_form"), names_to = "env", values_to = "env_value")%>%
    # left_join(sub_labels) %>%
    filter(trait_name %in% c("leaf_area","seed_dry_mass","leaf_capital_delta13C","leaf_mass_per_area","leaf_N_per_area_calc","wood_density","plant_height","huber_value")) %>%
    group_by(trait_name) %>%
    ungroup() %>%
    drop_na(value, env_value) %>%
    group_by(trait_name) %>%
    ungroup() -> data_for_plot
  
  data_for_plot$trait_name <- factor(data_for_plot$trait_name, labels = c("H[v]","LA","Delta^13~C~(`\u2030`)", "LMA~(mg~mm^{-2})", "N[area]~(mg~mm^{-2})", "H~(m)", "SM~(mg)", "Wood~density~(mg~mm^{-3})"))

data_for_plot %>%
  ggplot(aes(env_value, value,group = trait_name)) +
  geom_point(alpha = 0.5, aes(colour = growth_form)) + 
  scale_y_log10() +
  scale_x_log10(limits = c(min(woody_field_traits_georef_mean_climate$prec, na.rm = T), max(woody_field_traits_georef_mean_climate$prec, na.rm = T))) +
  theme_classic() +
  facet_wrap(~trait_name, scales = "free_y", 
                strip.position = "left",
               labeller = label_parsed) +
     theme(strip.background = element_blank(),
           strip.placement = "outside") +
  xlab("MAP (mm)") +
  ylab("") +
  theme(text=element_text(size=20), legend.position = "none") +
    # geom_text(mapping = aes(x = 0, y = Inf, label = label, group=env),
    #         hjust = "inward", vjust = "inward",
    #         inherit.aes = FALSE, check_overlap = TRUE, size = 6) +
  scale_color_manual(values = c("#009292","#db6d00")) +
  geom_smooth(method = "lm", col = "black") +
  tag_facets() ->p

  png("output/manuscript/gridded_relationships_prec_overall.png", height = 1500, width = 2666, res = 180)
p 
dev.off()

woody_field_traits_georef_mean_climate %>%
  drop_na(growth_form) %>%
    dplyr::select(trait_name, value, tavg, growth_form)%>%
    pivot_longer(-c("trait_name", "value","growth_form"), names_to = "env", values_to = "env_value")%>%
    # left_join(sub_labels) %>%
    filter(trait_name %in% c("leaf_area","seed_dry_mass","leaf_capital_delta13C","leaf_mass_per_area","leaf_N_per_area_calc","wood_density","plant_height","huber_value")) %>%
    group_by(trait_name) %>%
    ungroup() %>%
    drop_na(value, env_value) %>%
    group_by(trait_name) %>%
    ungroup() -> data_for_plot
  
  data_for_plot$trait_name <- factor(data_for_plot$trait_name, labels = c("H[v]","LA","Delta^13~C~(`\u2030`)", "LMA~(mg~mm^{-2})", "N[area]~(mg~mm^{-2})", "H~(m)", "SM~(mg)", "Wood~density~(mg~mm^{-3})"))

data_for_plot %>%
  ggplot(aes(env_value, value,group = trait_name)) +
  geom_point(alpha = 0.5, aes(colour = growth_form)) + 
  scale_y_log10() +
  scale_x_log10(limits = c(min(woody_field_traits_georef_mean_climate$tavg, na.rm = T), max(woody_field_traits_georef_mean_climate$tavg, na.rm = T))) +
  theme_classic() +
  facet_wrap(~trait_name, scales = "free_y", 
                strip.position = "left",
               labeller = label_parsed) +
     theme(strip.background = element_blank(),
           strip.placement = "outside") +
  xlab(expression("MAT ("*~degree*C*")")) +
  ylab("") +
  theme(text=element_text(size=20), legend.position = "none") +
    # geom_text(mapping = aes(x = 0, y = Inf, label = label, group=env),
    #         hjust = "inward", vjust = "inward",
    #         inherit.aes = FALSE, check_overlap = TRUE, size = 6) +
  scale_color_manual(values = c("#009292","#db6d00")) +
  geom_smooth(method = "lm", col = "black") +
  tag_facets() ->p


  png("output/manuscript/gridded_relationships_temp_overall.png", height = 1500, width = 2666, res = 180)
p 
dev.off()
```


```{r}
png("output/manuscript/supps_ellipse_woody.png", height=1300, width=2000, res=275)

woody_field_traits_georef_mean_climate_wide %>%
  filter(growth_form == "Woody") %>%
  dplyr::select(-c(taxon_name, growth_form, woody_climber)) -> for_corrplot
trait_env_correlation<- cor(for_corrplot, use="pairwise.complete.obs")
trait_env_correlation <- trait_env_correlation[c("log_seed_mass","leaf_capital_delta13C","log_leaf_area","log_LMA","log_wood_dens","log_leaf_N","log_height","log_huber"),c("log_Prec", "MI","AI_thorn","log_prec_wq","sqrt_prec_dq","Temp","GDD_5","Max_temp","log_VPD","SWdown","sqrt_elev","log_Prec_CV","temp_cv")]
colnames(trait_env_correlation) <- c("log(MAP)", "MI","AI_thorn","log_prec_wq","sqrt_prec_dq","MAT","GDD_5","max(Temp)","log(VPD)","SWDown","sqrt(Elevation)","log(PrecCV)","log(TempCV)")
rownames(trait_env_correlation) <- c("log(seed mass)","leaf capital delta13C","log(leaf area)","log(leaf mass per area)","log(wood density)","log(leaf nitrogen per area)","log(plant height)", "log(huber value)")
corrplot::corrplot(trait_env_correlation, method="ellipse")

dev.off()

png("output/manuscript/supps_ellipse_non_woody.png", height=1300, width=2000, res=275)

woody_field_traits_georef_mean_climate_wide %>%
  filter(growth_form == "Non-woody") %>%
  dplyr::select(-c(taxon_name, growth_form, woody_climber)) -> for_corrplot
trait_env_correlation<- cor(for_corrplot, use="pairwise.complete.obs")
trait_env_correlation <- trait_env_correlation[c("log_seed_mass","leaf_capital_delta13C","log_leaf_area","log_LMA","log_leaf_N","log_height"),c("log_Prec", "MI","AI_thorn","log_prec_wq","sqrt_prec_dq","Temp","GDD_5","Max_temp","log_VPD","SWdown","sqrt_elev","log_Prec_CV","temp_cv")]
colnames(trait_env_correlation) <- c("log(MAP)", "MI","AI_thorn","log_prec_wq","sqrt_prec_dq","MAT","GDD_5","max(Temp)","log(VPD)","SWDown","sqrt(Elevation)","log(PrecCV)","log(TempCV)")
rownames(trait_env_correlation) <- c("log(seed mass)","leaf capital delta13C","log(leaf area)","log(leaf mass per area)","log(leaf nitrogen per area)","log(plant height)")
corrplot::corrplot(trait_env_correlation, method="ellipse")

dev.off()

png("output/manuscript/supps_ellipse_overall.png", height=1300, width=2000, res=275)

woody_field_traits_georef_mean_climate_wide %>%
  filter(growth_form %in% c("Woody","Non-woody")) %>%
  dplyr::select(-c(taxon_name, growth_form, woody_climber)) -> for_corrplot
trait_env_correlation<- cor(for_corrplot, use="pairwise.complete.obs")
trait_env_correlation <- trait_env_correlation[c("log_seed_mass","leaf_capital_delta13C","log_leaf_area","log_LMA","log_leaf_N","log_height"),c("log_Prec", "MI","AI_thorn","log_prec_wq","sqrt_prec_dq","Temp","GDD_5","Max_temp","log_VPD","SWdown","sqrt_elev","log_Prec_CV","temp_cv")]
colnames(trait_env_correlation) <- c("log(MAP)", "MI","AI_thorn","log_prec_wq","sqrt_prec_dq","MAT","GDD_5","max(Temp)","log(VPD)","SWDown","sqrt(Elevation)","log(PrecCV)","log(TempCV)")
rownames(trait_env_correlation) <- c("log(seed mass)","leaf capital delta13C","log(leaf area)","log(leaf mass per area)","log(leaf nitrogen per area)","log(plant height)")
corrplot::corrplot(trait_env_correlation, method="ellipse")

dev.off()
```



#Estimate relative importance

```{r}
estimate_relative_importance <- function(data, woodiness_excluded, environmental_variables){
  data %>%
    filter(growth_form != woodiness_excluded) -> data
  
  function_call <- paste("value","~",paste(environmental_variables, collapse = "+"))
  fit <- lm(function_call, data)
  
  ci<-calc.relimp(fit, type = "lmg")
  # plot(ci)
  model_R2 = ci@R2
  model_lmg = ci@lmg
  return(tibble(model_R2, model_lmg, envir = unlist(environmental_variables)))
  # return(fit)
}
```


```{r}
woody_field_traits_georef_mean_climate_wide %>%
  dplyr::select(c("log_LMA", "log_leaf_capital_delta13C", "log_height", "log_seed_mass", "log_leaf_area", "log_leaf_N", "log_wood_dens", "log_huber", 'temp_cv', "log_Prec", "Temp", "VPD", "SWdown", "sqrt_elev","log_Prec_CV", "growth_form", 'woody_climber', "cell")) %>%
  mutate(scale_temp_cv = as.numeric(scale(temp_cv)), 
         scale_log_prec = as.numeric(scale(log_Prec)),
         scale_MAT = as.numeric(scale(Temp)),
         scale_VPD = as.numeric(scale(VPD)),
         scale_SRAD = as.numeric(scale(SWdown)),
         scale_elev = as.numeric(scale(sqrt_elev)),
         scale_log_Prec_CV = as.numeric(scale(log_Prec_CV))) %>%
  pivot_longer(cols = c(log_LMA, log_leaf_capital_delta13C, log_height, log_seed_mass, log_leaf_area, log_leaf_N, log_wood_dens, log_huber)) %>%
  group_by(name) %>%
  nest() %>%
  expand_grid(woodiness_excluded = c("Woody", "Non-woody",""))  %>%  
  filter(!(woodiness_excluded %in% c("", "Woody") & name %in% c("log_wood_dens","log_huber"))) %>%
  mutate(ci = map2(data, woodiness_excluded,estimate_relative_importance,list("scale_log_prec","scale_temp_cv","scale_MAT","scale_VPD","scale_SRAD","scale_elev","scale_log_Prec_CV"))) -> output

output %>% 
  unnest(ci) %>% 
  mutate(growth_form = case_when(
    woodiness_excluded == "Non-woody" ~ "Woody",
    woodiness_excluded == "Woody" ~ "Non-woody",
    woodiness_excluded == "" ~ "All taxa")
  ) %>%
  filter(!(growth_form %in% c("All taxa", "Non-woody") & name %in% c("log_wood_dens","log_huber"))) %>%
  mutate(name = factor(name,
                       levels = c("log_wood_dens","log_huber", "log_LMA", "log_leaf_capital_delta13C", "log_height", "log_seed_mass", "log_leaf_area", "log_leaf_N"),
                       labels = c("WD","H[v]","LMA","Delta^13~C", "H","SM","LA", "N[area]"))) %>%
  mutate(envir = factor(envir,
                       levels = c("scale_log_prec", "scale_MAT", "scale_VPD", "scale_SRAD", "scale_elev", "scale_log_Prec_CV","scale_temp_cv"),
                       labels = c("MAP","MAT","VPD","S[down]", "Elev","Prec[seas]", "Temp[seas]"))) %>%
  group_by(name, growth_form) %>%
  mutate(max_lmg = max(model_lmg)) %>%
  ungroup() %>%
  # mutate(envir = reorder(envir, model_lmg)) %>%
  mutate(growth_form = factor(growth_form, 
                       levels = c("Woody","Non-woody", "All taxa"))) -> data_for_plot


data_for_plot %>%
  ggplot() +
  geom_col(aes(x= model_lmg, y = name, fill = envir)) +
  facet_wrap(~growth_form) +
  theme_classic() +
  scale_y_discrete(labels=scales::label_parse()) +
  scale_fill_discrete(labels=scales::label_parse()) +
  ylab("") +
  xlab("Average proportion of variation explained") +
  theme(text = element_text(size = 18)) +
  theme(legend.text.align = 0) +
  labs(fill = "Environment")
```

Whittaker distribution of each trait

```{r}
png("output/manuscript/whittaker_plot.png", height=4000, width=4000, res=300)
purrr::map(c(core_traits, "legend"), austraits_climate_space) -> whittaker_plots
cowplot::plot_grid(plotlist = whittaker_plots)

dev.off()
```


```{r}

austraits_climate_space <- function(core_trait){
  
  aus_data = bind_cols(au_values_prec, au_values_temp)
  if(length(core_trait) == 1){
  if(core_trait == "legend"){
    
    ggplot() +
      geom_polygon(
        data = Whittaker_biomes,
        aes(x    = temp_c,
            y    = precp_cm,
            fill = biome),
        colour = "gray98",
        # colour of polygon border
        size   = 0.1
      ) +
      
      # set color for  the temperature - precipitation data points and the the AusTraits Sites
      scale_colour_manual(name = "Australian climate space", values = c("#FF7F50", "#233D4D")) +
      scale_fill_manual(
        name   = "Whittaker biomes",
        breaks = names(Ricklefs_colors),
        labels = names(Ricklefs_colors),
        values =  alpha(Ricklefs_colors, 0.5)
        
      ) +
      theme_classic() +
      guides(colour = guide_legend(override.aes = list(alpha = 1, size = 2))) +
      xlab(expression(Temperature (degree * C))) +
      ylab(" Precipitation (cm/yr)") +
      theme(text = element_text(size = 12))  +
      theme(
        legend.justification = c(-0.1, 0),
        legend.position = c(0.005, 0.25),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.background = element_rect(fill=NA)
      ) -> p
    
    p <- cowplot::get_legend(p)
  } else{
  
   woody_field_traits_georef_mean_climate_wide %>%
    drop_na(core_trait) -> site_data
  
  ggplot() +
  geom_polygon(
    data = Whittaker_biomes,
    aes(x    = temp_c,
        y    = precp_cm,
        fill = biome),
    colour = "gray98",
    # colour of polygon border
    size   = 0.1
  ) +
  
  # set color for  the temperature - precipitation data points and the the AusTraits Sites
  scale_colour_manual(name = "Australian climate space", values = c("#FF7F50", "#233D4D")) +
  scale_fill_manual(
    name   = "Whittaker biomes",
    breaks = names(Ricklefs_colors),
    labels = names(Ricklefs_colors),
    values =  alpha(Ricklefs_colors, 0.5)
    
  ) +
  theme_classic() +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 2))) +
  xlab(expression(Temperature (degree * C))) +
  ylab(" Precipitation (cm/yr)") +
  theme(text = element_text(size = 12))  +
  theme(
    legend.justification = c(-0.1, 0),
    legend.position = "none",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 10)
  ) +
  geom_point(
    data = aus_data,
    aes(x = temp,
        y = prec/10),
    col = "orange", alpha = 0.1
  ) +
  geom_point(
    data = site_data,
    aes(x = Temp,
        y = Prec/10)
    )+
    ggtitle(core_trait) ->p
  }
}
  else{
    woody_field_traits_georef_mean_climate_wide %>%
      filter(growth_form == "Woody") %>%
      pivot_longer(cols = c("leaf_mass_per_area")) %>%
      drop_na(value) %>%
      group_by(cells) %>%
      mutate(value = mean(value)) -> site_data
        ggplot() +
      geom_polygon(
        data = Whittaker_biomes,
        aes(x    = temp_c,
            y    = precp_cm*10,
            fill = biome),
        colour = "gray98",
        # colour of polygon border
        size   = 0.1
      ) +
      
      # set color for  the temperature - precipitation data points and the the AusTraits Sites
      scale_fill_manual(
        name   = "Whittaker biomes",
        breaks = names(Ricklefs_colors),
        labels = names(Ricklefs_colors),
        values =  alpha(Ricklefs_colors, 0.5)
        
      ) +
      theme_classic() +
      guides(colour = guide_legend(override.aes = list(alpha = 1, size = 2))) +
      xlab(expression(Mean~annual~temperature~(degree * C))) +
      ylab("Mean annual precipitation (mm)") +
      theme(text = element_text(size = 12))  +
      theme(
        legend.justification = c(-0.1, 0),
        legend.position=c(.005,.5),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 10)
      ) +
      geom_point(
        data = site_data,
        aes(x = Temp,
            y = Prec,
            colour = value) 
      ) +
          scale_color_continuous(trans = "log")->p 
  }
  
}
```

```{r}
png("output/manuscript/conceptual_figure_panel_a.png")

plot_data <- tibble(Trait =  seq(0, 1, by = 0.1),
       Environment = seq(0, 1, by = 0.1))
plot_data %>%
  ggplot(aes(x = Environment, y = Trait)) +
  theme_classic() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        text= element_text(size= 20))-> p
  
line_data_panel_a <- tibble(x = c(0,0,1,1), y = c(0.2, 0.5, 0.2, 0.8), line = c("lower","upper","lower","upper"))
polygon_data_panel_a <- tibble(x = c(0,0,1,1), y = c(0.2, 0.5, 0.8, 0.2), line = c("lower","upper","lower","upper"))

p +
  geom_line(data = line_data_panel_a, aes(x = x, y = y, group = line), linewidth= 2) +
  ylim(c(0,1)) +
  geom_polygon(data = polygon_data_panel_a, aes(x = x, y= y), fill = "blue", alpha = 0.4) +
  ylim(c(0.1,0.9))  -> a


line_data_panel_b <- tibble(x = c(0,1), y = c(0.1, 0.8))
UPPER_line_data_panel_b <- tibble(x = c(0,1), y = c(0.2, 0.9))
LOWER_line_data_panel_b <- tibble(x = c(0,1), y = c(0.0, 0.7))


points_data_panel_b <- tibble(x = seq(0.1,0.9, length.out = 5), y = 0.7*seq(0.1,0.9, length.out = 5)+0.1, z = seq(0,1,length.out = 5))

p +
  geom_line(data = line_data_panel_b, aes(x = x, y = y), linewidth= 2) +
  geom_line(data = UPPER_line_data_panel_b, aes(x = x, y = y), linewidth= 1, linetype = "dashed") +
  geom_line(data = LOWER_line_data_panel_b, aes(x = x, y = y), linewidth= 1, linetype = "dashed") +
  geom_point(data = points_data_panel_b , aes(x = x, y = y, colour = z), size = 7) +
  ylim(c(0,1)) +
  theme(legend.position = "none")+ 
  ylim(c(-0.1,1.1))  -> b




points_data_panel_c <- tibble(x = c(seq(0.1,0.9, length.out = 5),
                                    seq(0.1,0.9, length.out = 5),
                                    seq(0.1,0.9, length.out = 5)), 
                              y = c(0.8*seq(0.1,0.9, length.out = 5)+0.2,
                                    0.7*seq(0.1,0.9, length.out = 5)+0.15,
                                    0.7*seq(0.1,0.9, length.out = 5)), z = y)

line_data_panel_c <- tibble(x = c(0,1), y = c(0.125, 0.15 + 0.7*1))
UPPER_line_data_panel_c <- tibble(x = c(0,1), y = c(0.3, 0.3 + 0.8*1))
LOWER_line_data_panel_c <- tibble(x = c(0,1), y = c(-.05, 0.55))

p +
  geom_point(data = points_data_panel_c , aes(x = x, y = y, colour = z), size = 5) +
  ylim(c(0,1)) +
  theme(legend.position = "none") +
  geom_line(data = line_data_panel_c, aes(x = x, y = y), linewidth = 2) +
  geom_line(data = UPPER_line_data_panel_c, aes(x = x, y = y), linewidth = 1, linetype = "dashed") + 
  geom_line(data = LOWER_line_data_panel_c, aes(x = x, y = y), linewidth = 1, linetype = "dashed") + 
  ylim(c(-0.1,1.1)) -> c


png("output/manuscript/theoretical_plots_a.png", width = 800, height = 800, res = 200)
a
dev.off()


png("output/manuscript/theoretical_plots_b.png", width = 800, height = 800, res = 200)
b
dev.off()

png("output/manuscript/theoretical_plots_c.png", width = 800, height = 800, res = 200)
c
dev.off()



png("output/manuscript/theoretical_plots.png", width = 2000, height = 1000, res = 200)
cowplot::plot_grid(a,b,c, nrow = 1)
dev.off()

```


```{r}
estimate_prop_of_possible_variation <- function(...){
  
  data <- tibble(...)
  data %>%
    unnest() %>%
    filter(growth_form != woodiness_excluded) -> data
  
  function_call <- paste("trait_value","~","env_value", "+ (1|cells)")
  fit <- lmer(function_call, data)
  r2 <- MuMIn::r.squaredGLMM(fit)

  return(tibble(marg_r2 = r2[1], cond_r2 = r2[2]))
  # return(fit)
}
```

```{r}
woody_field_traits_georef_mean_climate_wide %>%
  dplyr::select(growth_form, log_leaf_capital_delta13C, log_height, log_leaf_N, log_LMA, log_seed_mass, log_leaf_area, log_wood_dens, log_huber, log_Prec, Temp, GDD_5, SWdown, Max_temp, elev, log_VPD, climaticMoistureIndex, aridityIndexThornthwaite, log_MI,log_Prec_CV,temp_cv, cells) %>%
  pivot_longer(cols = c(log_leaf_capital_delta13C, 
                        log_leaf_area,
                        log_seed_mass,
                        log_LMA,
                        log_leaf_N,
                        log_height,
                        log_wood_dens,
                        log_huber),
                        names_to = "trait_name", values_to = "trait_value") %>%
  pivot_longer(cols = c(log_Prec, Temp, GDD_5, SWdown, Max_temp, elev, log_VPD, climaticMoistureIndex, aridityIndexThornthwaite, log_MI,log_Prec_CV,temp_cv), names_to = "env_name", values_to = "env_value") %>% 
  group_by(trait_name, env_name) %>%
  nest() %>%
  expand_grid(woodiness_excluded = c("Woody", "Non-woody", "")) %>%
  mutate(r2 = pmap(., estimate_prop_of_possible_variation))  -> output

output %>% dplyr::select(-data) %>% unnest(r2) %>% View()

output %>% dplyr::select(-data) %>% unnest(r2) %>% mutate(r2_prop = marg_r2/cond_r2)%>% filter(woodiness_excluded == "Non-woody") %>% dplyr::select(-marg_r2, -r2_prop) %>% pivot_wider(names_from = "env_name", values_from = cond_r2) %>% View()

```




```{r}
library(lme4)
library("MuMIn")
library(nlme)

MuMIn::r.squaredGLMM(lmer(log_leaf_capital_delta13C~1 + (1|cells), woody_field_traits_georef_mean_climate_wide %>% filter(growth_form == "Woody"))) 

MuMIn::r.squaredGLMM(lmer(log_wood_dens~log_Prec + (1|cells), woody_field_traits_georef_mean_climate_wide %>% filter(growth_form == "Woody"))) 


fit1 <- lme(log_LMA~1, random = (~1|cells), woody_field_traits_georef_mean_climate_wide %>% filter(growth_form == "Woody"), na.action = na.omit)

summary(lmer(log_LMA~1 + (1|cells), 
             woody_field_traits_georef_mean_climate_wide)) -> fit_null

r.squared(fit_null)

```




#Estimate relative importance

```{r}
estimate_relative_importance <- function(data, woodiness_excluded, environmental_variables){
  data %>%
    filter(growth_form != woodiness_excluded) -> data
  
  function_call <- paste("value","~",paste(environmental_variables[[2]], collapse = "+"))
  fit <- lm(function_call, data)
  p1 <- visreg::visreg(fit, "scale_temp_cv")

  
  function_call <- paste("value","~",paste(environmental_variables, collapse = "+"))
  fit <- lm(function_call, data)
  p2 <- visreg::visreg(fit, "scale_temp_cv")

  return(list(p1, p2))
  # return(fit)
}
```


```{r}
woody_field_traits_georef_mean_climate_wide %>%
  dplyr::select(c("log_LMA", "log_leaf_capital_delta13C", "log_height", "log_seed_mass", "log_leaf_area", "log_leaf_N", "log_wood_dens", "log_huber", 'temp_cv', "log_Prec", "Temp", "VPD", "SWdown", "sqrt_elev","log_Prec_CV", "growth_form", 'woody_climber', "cell")) %>%
  mutate(scale_temp_cv = as.numeric(scale(temp_cv)), 
         scale_log_prec = as.numeric(scale(log_Prec)),
         scale_MAT = as.numeric(scale(Temp)),
         scale_VPD = as.numeric(scale(VPD)),
         scale_SRAD = as.numeric(scale(SWdown)),
         scale_elev = as.numeric(scale(sqrt_elev)),
         scale_log_Prec_CV = as.numeric(scale(log_Prec_CV))) %>%
  pivot_longer(cols = c(log_LMA, log_leaf_capital_delta13C, log_height, log_seed_mass, log_leaf_area, log_leaf_N, log_wood_dens, log_huber)) %>%
  mutate(trait_name = name) %>%
  group_by(name) %>%
  nest() %>%
  expand_grid(woodiness_excluded = c("Non-woody"))  %>%  
  filter(!(woodiness_excluded %in% c("", "Woody") & name %in% c("log_wood_dens","log_huber"))) %>%
  mutate(ci = map2(data, woodiness_excluded,estimate_relative_importance,list("scale_log_prec","scale_temp_cv"))) -> output

output %>% 
  unnest(ci) %>% 
  mutate(growth_form = case_when(
    woodiness_excluded == "Non-woody" ~ "Woody",
    woodiness_excluded == "Woody" ~ "Non-woody",
    woodiness_excluded == "" ~ "All taxa")
  ) %>%
  filter(!(growth_form %in% c("All taxa", "Non-woody") & name %in% c("log_wood_dens","log_huber"))) %>%
  mutate(name = factor(name,
                       levels = c("log_wood_dens","log_huber", "log_LMA", "log_leaf_capital_delta13C", "log_height", "log_seed_mass", "log_leaf_area", "log_leaf_N"),
                       labels = c("WD","H[v]","LMA","Delta^13~C", "H","SM","LA", "N[area]"))) %>%
  mutate(envir = factor(envir,
                       levels = c("scale_log_prec", "scale_MAT", "scale_VPD", "scale_SRAD", "scale_elev", "scale_log_Prec_CV","scale_temp_cv"),
                       labels = c("MAP","MAT","VPD","S[down]", "Elev","Prec[seas]", "Temp[seas]"))) %>%
  group_by(name, growth_form) %>%
  mutate(max_lmg = max(model_lmg)) %>%
  ungroup() %>%
  # mutate(envir = reorder(envir, model_lmg)) %>%
  mutate(growth_form = factor(growth_form, 
                       levels = c("Woody","Non-woody", "All taxa"))) -> data_for_plot


data_for_plot %>%
  ggplot() +
  geom_col(aes(x= model_lmg, y = name, fill = envir)) +
  facet_wrap(~growth_form) +
  theme_classic() +
  scale_y_discrete(labels=scales::label_parse()) +
  scale_fill_discrete(labels=scales::label_parse()) +
  ylab("") +
  xlab("Average proportion of variation explained") +
  theme(text = element_text(size = 18)) +
  theme(legend.text.align = 0) +
  labs(fill = "Environment")
```

