---
title: Report on data in AusTraits
output:
  html_document:
    df_print: kable
    highlight: tango
    keep_md: no
    smart: no
    theme: yeti
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options:
  chunk_output_type: console
---

Load data
<<<<<<< Updated upstream
```{r, warning=FALSE, message=FALSE}
=======
```{r include=FALSE}
>>>>>>> Stashed changes
# Install dependencies as needed
# devtools::install_deps()
# Load packages and functions for this analysis
devtools::load_all()
source("R/load_climate_data.R")

# load data
austraits <- readRDS("data/austraits-3.0.2.rds")

au_bioclim <- load_climate_data()

# match with climate data
au_sites_clim <-
  austraits$sites %>%
  filter(site_property %in% c("longitude (deg)", "latitude (deg)")) %>%
  spread(site_property, value) %>%
  rowid_to_column("ID") %>%
  rename(latitude = `latitude (deg)`, longitude = `longitude (deg)`) %>%
  mutate_at(c("latitude", "longitude"), ~suppressWarnings(as.numeric(.x))) %>%
  combine_occurence_climate(au_bioclim)
<<<<<<< Updated upstream
=======

raster::plot(au_bioclim[[3]])

na_sites<-au_sites_clim %>% filter_all(any_vars(is.na(.)))
```

```{r, fig.height=10, fig.width=14}

raster::plot(au_bioclim[[1]], col=rev(terrain.colors(1000)))

raster::plot(au_bioclim[[1]], col=rev(terrain.colors(1000)))
points(au_sites_clim$lat~au_sites_clim$lon, pch=19, col=rev(terrain.colors(1000))[round((au_sites_clim$Temp/max(au_sites_clim$Temp, na.rm=T)*1000-160)*1.19,0)])


raster::plot(au_bioclim[[2]], col=rev(terrain.colors(100)))
points(au_sites_clim$lat~au_sites_clim$lon, pch=19, col=rev(terrain.colors(100))[round(au_sites_clim$Prec/max(au_sites_clim$Prec, na.rm=T)*100,0)])


>>>>>>> Stashed changes
```

Calculate leaf_N_per_area and leaf_area for additional observations:
```{r include=FALSE}
#two current problems, in summarise argument, `first` should be `mean` but breaking on instances with a single observation (almost all). Second, this code means the newly calculated values don't have value_type and replicate linked to them

austraits$traits %>% 
  filter(trait_name %in% c("leaf_area","leaf_dry_mass","specific_leaf_area","leaf_N_per_dry_mass",
                           "leaf_N_per_area","leaf_P_per_dry_mass","leaf_P_per_area")) %>%
  select(trait_name, value, observation_id, taxon_name,dataset_id,site_name, context_name, date) %>%
  mutate(value = as.numeric(value)) %>%
  group_by(trait_name, observation_id, taxon_name, dataset_id, site_name, context_name, date) %>%
  summarise(value = mean(value)) %>%
  ungroup() %>%
  spread(key = trait_name, value = value) %>% 
  mutate(specific_leaf_area = as.numeric(specific_leaf_area),
         leaf_N_per_dry_mass = as.numeric(leaf_N_per_dry_mass),
         leaf_area = as.numeric(leaf_area),
         leaf_dry_mass = as.numeric(leaf_dry_mass),
         leaf_N_per_area = as.numeric(leaf_N_per_area)) %>%
  mutate(leaf_N_calc_SLA = leaf_N_per_dry_mass / specific_leaf_area,
         leaf_N_calc_LA = leaf_N_per_dry_mass*leaf_dry_mass/leaf_area,
         leaf_N_per_area_combined = ifelse(!is.na(leaf_N_per_area),leaf_N_per_area,leaf_N_calc_SLA),
         leaf_N_per_area_combined = ifelse(!is.na(leaf_N_per_area_combined),leaf_N_per_area_combined,leaf_N_calc_LA),
         specific_leaf_area_combined = ifelse(!is.na(specific_leaf_area),specific_leaf_area,leaf_area/leaf_dry_mass),
         leaf_area_combined = ifelse(!is.na(leaf_area),leaf_area,specific_leaf_area*leaf_dry_mass)) %>%
  mutate(leaf_P_calc_SLA = leaf_P_per_dry_mass / specific_leaf_area,
         leaf_P_calc_LA = leaf_P_per_dry_mass*leaf_dry_mass/leaf_area,
         leaf_P_per_area_combined = ifelse(!is.na(leaf_P_per_area),leaf_P_per_area,leaf_P_calc_SLA),
         leaf_P_per_area_combined = ifelse(!is.na(leaf_P_per_area_combined),leaf_P_per_area_combined,leaf_P_calc_LA))->calculated_data

calculated_data %>%
  select(observation_id, taxon_name, dataset_id, site_name, context_name, date, 
         leaf_P_per_area_combined) %>%
  filter(!is.na(leaf_P_per_area_combined)) %>%
  mutate(units = "g/m2",
         trait_name = "leaf_P_per_area_calc",
         value = as.character(leaf_P_per_area_combined)) %>%
  select(-leaf_P_per_area_combined) -> leaf_P_per_area_calcuated

calculated_data %>%
  select(observation_id, taxon_name, dataset_id, site_name, context_name, date, 
         leaf_N_per_area_combined) %>%
  filter(!is.na(leaf_N_per_area_combined)) %>%
  mutate(units = "g/m2",
         trait_name = "leaf_N_per_area_calc",
         value = as.character(leaf_N_per_area_combined)) %>%
  select(-leaf_N_per_area_combined) -> leaf_N_per_area_calcuated

calculated_data %>%
  select(observation_id, taxon_name, dataset_id, site_name, context_name, date, 
         leaf_area_combined) %>%
  filter(!is.na(leaf_area_combined)) %>%
  mutate(units = "mm2",
         trait_name = "leaf_area_calc",
         value = as.character(leaf_area_combined)) %>%
  select(-leaf_area_combined) -> leaf_area_calcuated

calculated_data %>%
  select(observation_id, taxon_name, dataset_id, site_name, context_name, date, 
         specific_leaf_area_combined) %>%
  filter(!is.na(specific_leaf_area_combined)) %>%
  mutate(units = "mm2/mg",
         trait_name = "specific_leaf_area_calc",
         value = as.character(specific_leaf_area_combined)) %>%
  select(-specific_leaf_area_combined) -> specific_leaf_area_calcuated


```

<<<<<<< Updated upstream
Select subset of traits:
```{r}
# add max height, N_area, delta13_c
# Also do for some alternatives: e.g. leaf_width, leaf_area ?
trait_names_short <- c("leaf_width","leaf_delta13C","plant_height","leaf_N_per_area","specific_leaf_area","leaf_area", "wood_density", "seed_mass", "plant_growth_form")
trait_names <- c("leaf_N_per_area","leaf_N_per_area_calc","specific_leaf_area","specific_leaf_area_calc","leaf_area","leaf_area_calc","leaf_width","leaf_delta13C","wood_density","seed_mass","plant_height", "plant_growth_form")

traits <- 
  austraits$traits %>%
  filter(trait_name %in% trait_names_short) %>% 
=======
Extract values for delta13c from ci:ca ratios using equations in Farqhuar (1989) and Guerrieri et al. (2019; PNAS)

The ci:ca ratio is linked to the carbon isotope fraction of plant samples according the following equation:

$$a+(b-a)\frac{c_i}{c_a}=\frac{\delta^{13}C_a-\delta^{13}C_c}{1+\frac{\delta^{13}C_c}{1000}},$$
where $\delta^{13}C_a$ is approximately equal to -8, a = 4.4 and b = 27.Rearranging the above equation to solve $\delta^{13}C_c$ gives:

$$\delta^{13}C_c = \frac{-8000-1000*a+(b-a)\frac{c_i}{c_a}}{a+(b-a)\frac{c_i}{c_a}+1000}$$


``` {r}
austraits$traits %>% 
  filter(trait_name %in% c("ci_over_ca", "leaf_delta13C")) %>%
  select(trait_name, value, observation_id, taxon_name,dataset_id,site_name, context_name, date) %>%
  mutate(value = as.numeric(value)) %>%
  group_by(trait_name, observation_id, taxon_name, dataset_id, site_name, context_name, date) %>%
  summarise(value = mean(value)) %>%
  ungroup() %>%
  spread(key = trait_name, value = value)%>%
  mutate(leaf_delta13C_calc = (-8000-1000*(4.4+(27-4.4)*ci_over_ca))/(4.4+(27-4.4)*ci_over_ca+1000))%>%
  mutate(leaf_delta13C_combined = ifelse(!is.na(leaf_delta13C),leaf_delta13C,leaf_delta13C_calc)) ->calculated_data

calculated_data %>%
  select(observation_id, taxon_name, dataset_id, site_name, context_name, date, 
         leaf_delta13C_combined) %>%
  filter(!is.na(leaf_delta13C_combined)) %>%
  mutate(units = "per mille",
         trait_name = "leaf_delta13C_calc",
         value = as.character(leaf_delta13C_combined)) %>%
  select(-leaf_delta13C_combined) -> leaf_delta13C_calculated



```

Select subset of traits:
```{r include=FALSE}

traits <- 
  austraits$traits %>%
>>>>>>> Stashed changes
  bind_rows(leaf_N_per_area_calcuated) %>%
  bind_rows(leaf_P_per_area_calcuated) %>%
  bind_rows(specific_leaf_area_calcuated) %>%
  bind_rows(leaf_delta13C_calculated)%>%
  bind_rows(leaf_area_calcuated) %>%
<<<<<<< Updated upstream
  left_join(by = c("dataset_id", "site_name"), au_sites_clim)
=======
  mutate(value = as.numeric(value)) %>%
  group_by(dataset_id, taxon_name, site_name, trait_name) %>%
  summarise(value=median(value)) %>%
  left_join(by = c("dataset_id", "site_name"), au_sites_clim)

>>>>>>> Stashed changes
```

Inspect data availability across growth forms (woody vs. non-woody)

<<<<<<< Updated upstream
```{r}
plant_growth_form_index <- 
  traits %>%
  filter(trait_name == "plant_growth_form") %>%
  rename("plant_growth_form" = "value") %>%
  mutate(plant_form_broad = ifelse(plant_growth_form %in% c("grass-tree","herb shrub","herb subshrub","herb_large shrub","prostate_shrub","shrub subshrub","shrub tree","shrub treelet","subshrub","shrub","treelet","tree","tree tree"), "woody", "non_woody")) 

# how many species for which growth form was recorded have both woody and non-woody forms
number_of_forms_species <- plant_growth_form_index %>% 
  group_by(taxon_name) %>%
  summarise(length(unique(plant_form_broad))) %>%
  rename(number_of_forms = 2)

# join number of growth forms to plant_growth_form_index
plant_growth_form_index <- plant_growth_form_index %>%
  left_join(number_of_forms_species) %>%
  select("taxon_name", "plant_form_broad", "plant_growth_form", "number_of_forms") %>%
  group_by(taxon_name) %>%
  slice(1)

traits <- traits %>%
  filter(!trait_name=="plant_growth_form") %>%
  mutate_at("value", as.numeric)

# reintroduce plant_growth_form index to trait dataframe
traits <- traits %>%
  left_join(plant_growth_form_index) 

# how are samples distributed across species with 1, 2 or no recorded broad growth form (i.e. woody or non-woody)
table(traits$number_of_forms, useNA = "ifany")
=======
```{r include=FALSE}
field_traits <- traits %>%
  filter(!(dataset_id %in% c("Ahrens_2019", "Crous_2013", "Crous_2019", "Du_2018",
                            "Du_2019", "Duan_2015", "EsperonRodriguez_2019",
                            "EsperonRodriguez_2020","Everingham_2020",
                            "Farrell_2012", "Farrell_2013", "Farrell_2017", "Firn_2019",
                            "Gallagher_2011_1", "Gallagher_2011_3", "Gallagher_2015", 
                            "Gardiner_2019", "Geange_2017", "Geange_2020", "Ghannoum_2010",
                            "Harrison_2009",
                            "Hassiotou_2009", "Huang_2015", "Jagdish_2020", "Lewis_2015",
                            "Lusk_2012", "Lusk_2014", "Moore_2019_2", "Reynolds_2018",
                            "Roderick_2002", "Schulze_2006", "Smith_2012",
                            "Tomlinson_2013","Tomlinson_2019", "Turner_2010", "Vesk_2019",
                            "Vlasveld_2018", "Warren_2005", "Warren_2006"))) %>%
  filter(!(site_name %in% c("Currency Creek Arboretum","Currency Creek Arboretum, South Australia",
                            "Australian National University glasshouse","MU_campus",
                            "Australian National Botanic Gardens","Australian National University"))) 
>>>>>>> Stashed changes
```

54600 trait samples belong to taxa which have been recorded to have  both woody and non-woody growth forms; 5555 trait samples belong to taxa which have not yet had a growth form recorded. Note that growth form is often not available for all trait observations of a given taxa. Will need to address how species with woody and non-woody forms should be addressed.

<<<<<<< Updated upstream
Create trait data frame for woody taxa only

```{r}
woody_traits <- traits %>%
# remove species which had two forms - will keep ones that had no growth form identified (i.e. NAs)
  filter(is.na(number_of_forms) | number_of_forms == 1 )  %>%
# finally, remove species which were exclusively non-woody
  filter(is.na(plant_form_broad) | plant_form_broad == "woody")
```

Create a new trait dataframe to only include observations with associated latitude and longitude
```{r}
traits_georef <- 
  traits %>%
  drop_na(latitude,longitude)
=======
``` {r include=FALSE}


austraits$traits %>%
  filter(trait_name == "plant_growth_form") %>%
  filter(value %in% c("hemi-epiphyte_shrub", "hemi-epiphyte_tree", "leafless_shrub_tree", "parasite_woody", 
                       "subshrub", "shrub", "treelet", "tree", "shrub tree", "shrub treelet",
                      "shrub subshrub","herb subshrub", "herb_large shrub", "herb shrub", "tree tree")) %>%
  distinct(taxon_name) -> woody_taxa
austraits$traits %>% filter(trait_name == "woodiness") %>% 
  filter(value %in% c("woody","semi_woody")) %>%
  distinct(taxon_name) %>%
    bind_rows(woody_taxa) %>%
    distinct(taxon_name) -> woody_taxa  
  
read_csv("data/no_woody_data.csv") %>%
  filter(woodiness %in% c("woody","semi_woody")) %>%
           distinct(taxon_name) %>%
           bind_rows(woody_taxa) %>%
           distinct(taxon_name) -> woody_taxa


woody_traits <- traits %>%
  filter(taxon_name %in% woody_taxa$taxon_name)

woody_field_traits <- field_traits %>%
  filter(taxon_name %in% woody_taxa$taxon_name)


```

Create a new trait dataframe to only include observations with associated latitude and longitude
```{r include=FALSE}
>>>>>>> Stashed changes

woody_traits_georef <- 
  woody_traits %>%
  drop_na(latitude,longitude)

traits_site_summary <- 
  traits %>% 
  group_by(dataset_id, trait_name, latitude, longitude, Temp, Prec, Prec_CV) %>% 
  summarise(nspp=n_distinct(taxon_name))

woody_traits_site_summary <- 
  traits %>% 
  group_by(dataset_id, trait_name, latitude, longitude, Temp, Prec, Prec_CV) %>% 
  summarise(nspp=n_distinct(taxon_name))
```

Write some summaries about availability of data.....


Plots:

Austraits sites across biomes - total dataset vs. woody only dataset

<<<<<<< Updated upstream
```{fig.height=12, fig.width=6}
austraits_climate_space <-
  plot_climate_space(traits, au_bioclim)
=======
woody_field_traits_georef <- 
  woody_field_traits %>%
  drop_na(lat,lon)
```
>>>>>>> Stashed changes

austraits_climate_space_woody <-
  plot_climate_space(woody_traits, au_bioclim)

<<<<<<< Updated upstream
austraits_climate_space / austraits_climate_space_woody
```

Woody trait dataset

```{fig.height=12, fig.width=6}
austraits_site_by_trait <-
  plot_site_map_by_trait(traits)

austraits_site_by_trait_woody <-
  plot_site_map_by_trait(woody_traits)

austraits_site_by_trait / austraits_site_by_trait_woody
```

Shading of points (blue -> yellow) indicates the number of neighbouring points (e.g. there is substantial seed mass data in SE Australia)

Compare the number of observations and taxa in the **total dataset** for each trait which are **georeferenced vs. not georeferenced**.

```{r}
number_observations_per_trait<-traits %>%
  group_by(trait_name) %>%
  summarise(n_total_obs=length(value), n_total_taxa=length(unique(taxon_name)))

number_observations_per_trait_georeferenced<-traits_georef %>%
  group_by(trait_name) %>%
  summarise(n_georef_obs=length(value), n_georef_taxa=length(unique(taxon_name))) %>%
  left_join(number_observations_per_trait)

number_observations_per_trait_georeferenced %>% 
  relocate(n_total_obs, .after=trait_name) %>%
  relocate(n_total_taxa, .before=n_georef_taxa) %>%
  kbl(caption = "Table 1: Total trait dataset. n_total_obs: Total number of observations of a given trait, n_georef_obs: Number of observations of a given trait with georeference, n_total_taxa: Total number of taxa with at least one observation of a given trait, n_georef_taxa: Number of taxa with at least one georeferenced observation of a given trait") %>%
  kable_classic(full_width = T, html_font = "Cambria")  

```

Compare the number of observations and taxa in the **woody dataset** for each trait which are **georeferenced vs. not georeferenced**.

```{r}
number_observations_per_trait<-woody_traits %>%
  group_by(trait_name) %>%
  summarise(n_total_obs=length(value), n_total_taxa=length(unique(taxon_name)))

number_observations_per_trait_georeferenced<-woody_traits_georef %>%
  group_by(trait_name) %>%
  summarise(n_georef_obs=length(value), n_georef_taxa=length(unique(taxon_name))) %>%
  left_join(number_observations_per_trait)

number_observations_per_trait_georeferenced %>% 
  relocate(n_total_obs, .after=trait_name) %>%
  relocate(n_total_taxa, .before=n_georef_taxa) %>%
  kbl(caption = "Table 2: Woody trait dataset. n_total_obs: Total number of observations of a given trait, n_georef_obs: Number of observations of a given trait with georeference, n_total_taxa: Total number of taxa with at least one observation of a given trait, n_georef_taxa: Number of taxa with at least one georeferenced observation of a given trait") %>%
  kable_classic(full_width = T, html_font = "Cambria")  
=======
Transform negative value traits to positive to facilitate plotting on transformed scales
```{r include=FALSE}
woody_field_traits_georef <- woody_field_traits_georef %>%
  #multiply leaf_delta13C by -1 to put it on the positive scale for log scaling
  mutate(value = if_else(trait_name=="water_potential_50percent_lost_conductivity", value*-1, value), trait_name = if_else(trait_name=="water_potential_50percent_lost_conductivity", paste0("-1 * ",trait_name, sep=""), trait_name)) %>%
  mutate(value = if_else(trait_name=="water_potential_88percent_lost_conductivity", value*-1, value), trait_name = if_else(trait_name=="water_potential_88percent_lost_conductivity", paste0("-1 * ",trait_name, sep=""), trait_name)) %>%
  mutate(value = if_else(trait_name=="leaf_delta13C", value*-1, value), trait_name = if_else(trait_name=="leaf_delta13C", paste0("-1 * ",trait_name, sep=""), trait_name)) %>%
  mutate(value = if_else(trait_name=="leaf_delta13C_calc", value*-1, value), trait_name = if_else(trait_name=="leaf_delta13C_calc", paste0("-1 * ",trait_name, sep=""), trait_name))
```

Define some core traits to focus on

```{r}

core_traits <- c("leaf_area_calc","-1 * leaf_delta13C","-1 * leaf_delta13C_calc","leaf_N_per_area_calc","leaf_P_per_area_calc","specific_leaf_area_calc","wood_density","plant_height", "seed_mass")
```

Distribution of sites across the climate space of Australia

```{r}
plot_site_map_by_trait(woody_field_traits_georef %>% filter(trait_name %in% core_traits))

raster::values(au_bioclim) %>% 
  as_data_frame() %>% 
  drop_na() %>%
  ggplot() +
  geom_hex(aes(x = Temp, y = Prec), colour="grey", fill="grey") +
  geom_hex(data = au_sites_clim, aes(x=Temp, y=Prec), show.legend = TRUE,inherit.aes = FALSE) +
  viridis::scale_fill_viridis(trans = "log10",limits=c(1,1000)) + 
  scale_y_log10()
    
raster::values(au_bioclim) %>% 
  as_data_frame() %>% 
  drop_na() %>%
  ggplot() +
  geom_hex(aes(x = Prec_CV, y = Prec), colour="grey", fill="grey") +
  geom_hex(data = au_sites_clim, aes(x = Prec_CV, y = Prec), show.legend = TRUE,inherit.aes = FALSE) +
  viridis::scale_fill_viridis(trans = "log10",limits=c(1,1000)) + 
  scale_y_log10()

raster::values(au_bioclim) %>% 
  as_data_frame() %>% 
  drop_na() %>%
  ggplot() +
  geom_hex(aes(x = Prec_CV, y = Temp), colour="grey", fill="grey") +
  geom_hex(data = au_sites_clim, aes(x = Prec_CV, y = Temp), show.legend = TRUE,inherit.aes = FALSE) +
  viridis::scale_fill_viridis(trans = "log10",limits=c(1,1000)) + 
  scale_y_log10()
>>>>>>> Stashed changes

```

Basic patterns with environment - all data

```{r, fig.width=10}
traits %>%
  #multiply leaf_delta13C by -1 to put it on the positive scale for log scaling
  mutate(value = if_else(trait_name=="leaf_delta13C", value*-1, value), trait_name = if_else(trait_name=="leaf_delta13C", paste0("-1 * ",trait_name, sep=""), trait_name)) %>%
    ggplot(aes(Prec, value, colour=plant_form_broad)) + geom_point(alpha = 0.2, stroke=0) + scale_x_log10() + scale_y_log10() + facet_wrap(~trait_name, scales = "free_y")
```

```{r, fig.width=10}
traits %>%
  #multiply leaf_delta13C by -1 to put it on the positive scale for log scaling
  mutate(value = if_else(trait_name=="leaf_delta13C", value*-1, value), trait_name = if_else(trait_name=="leaf_delta13C", paste0("-1 * ",trait_name, sep=""), trait_name)) %>%
  ggplot(aes(Temp, value, colour=plant_form_broad)) + geom_point(alpha = 0.2, stroke=0) + scale_x_log10() + scale_y_log10() + facet_wrap(~trait_name, scales = "free_y")
```

```{r, fig.width=10}
traits %>%
  #multiply leaf_delta13C by -1 to put it on the positive scale for log scaling
  mutate(value = if_else(trait_name=="leaf_delta13C", value*-1, value), trait_name = if_else(trait_name=="leaf_delta13C", paste0("-1 * ",trait_name, sep=""), trait_name)) %>%
  ggplot(aes(Prec_CV, value, colour=plant_form_broad)) + geom_point(alpha = 0.2, stroke=0) + scale_x_log10() + scale_y_log10() + facet_wrap(~trait_name, scales = "free_y")
```

Plot the distribution of these species across the climate space of Australia

```{r eval=FALSE, include=FALSE}
raster::values(au_bioclim) %>% 
  as_data_frame() %>% 
  drop_na() %>%
  ggplot() +
  geom_hex(aes(x = Temp, y = Prec), colour="grey", fill="grey") +
  geom_hex(data = woody_field_traits_georef_euc_sites, aes(x=Temp, y=Prec), show.legend = TRUE,inherit.aes = FALSE) +
  viridis::scale_fill_viridis(trans = "log10",limits=c(1,1000)) + 
  scale_y_log10()
    
raster::values(au_bioclim) %>% 
  as_data_frame() %>% 
  drop_na() %>%
  ggplot() +
  geom_hex(aes(x = Prec_CV, y = Prec), colour="grey", fill="grey") +
  geom_hex(data = woody_field_traits_georef_euc_sites, aes(x = Prec_CV, y = Prec), show.legend = TRUE,inherit.aes = FALSE) +
  viridis::scale_fill_viridis(trans = "log10",limits=c(1,1000)) + 
  scale_y_log10()

raster::values(au_bioclim) %>% 
  as_data_frame() %>% 
  drop_na() %>%
  ggplot() +
  geom_hex(aes(x = Prec_CV, y = Temp), colour="grey", fill="grey") +
  geom_hex(data = woody_field_traits_georef_euc_sites, aes(x = Prec_CV, y = Temp), show.legend = TRUE,inherit.aes = FALSE) +
  viridis::scale_fill_viridis(trans = "log10",limits=c(1,1000)) + 
  scale_y_log10()
```

Basic patterns with environment - woody only

```{r, fig.width=10}
<<<<<<< Updated upstream
woody_traits %>%
  #multiply leaf_delta13C by -1 to put it on the positive scale for log scaling
  mutate(value = if_else(trait_name=="leaf_delta13C", value*-1, value), trait_name = if_else(trait_name=="leaf_delta13C", paste0("-1 * ",trait_name, sep=""), trait_name)) %>%
    ggplot(aes(Prec, value)) + geom_point(alpha = 0.2, stroke=0) + scale_x_log10() + scale_y_log10() + facet_wrap(~trait_name, scales = "free_y")
```

```{r, fig.width=10}
woody_traits %>%
  #multiply leaf_delta13C by -1 to put it on the positive scale for log scaling
  mutate(value = if_else(trait_name=="leaf_delta13C", value*-1, value), trait_name = if_else(trait_name=="leaf_delta13C", paste0("-1 * ",trait_name, sep=""), trait_name)) %>%
  ggplot(aes(Temp, value)) + geom_point(alpha = 0.2, stroke=0) + scale_x_log10() + scale_y_log10() + facet_wrap(~trait_name, scales = "free_y")
```

```{r, fig.width=10}
woody_traits %>%
  #multiply leaf_delta13C by -1 to put it on the positive scale for log scaling
  mutate(value = if_else(trait_name=="leaf_delta13C", value*-1, value), trait_name = if_else(trait_name=="leaf_delta13C", paste0("-1 * ",trait_name, sep=""), trait_name)) %>%
  ggplot(aes(Prec_CV, value)) + geom_point(alpha = 0.2, stroke=0) + scale_x_log10() + scale_y_log10() + facet_wrap(~trait_name, scales = "free_y")
=======

woody_field_traits_georef %>%
  filter(trait_name %in% core_traits) %>%
  ggplot(aes(Prec, value)) + geom_point(alpha = 0.2, stroke=0)+scale_x_log10() + scale_y_log10() + facet_wrap(~trait_name, scales = "free_y")
```

```{r, fig.width=10}
woody_field_traits_georef %>%
  filter(trait_name %in% core_traits) %>%
  ggplot(aes(Temp, value)) + geom_point(alpha = 0.2, stroke=0)+scale_x_log10() + scale_y_log10() + facet_wrap(~trait_name, scales = "free_y")
```

```{r, fig.width=10}
woody_field_traits_georef %>%
  filter(trait_name %in% core_traits) %>%
  ggplot(aes(DNI, value)) + geom_point(alpha = 0.2, stroke=0)+ scale_y_log10() + facet_wrap(~trait_name, scales = "free_y")
>>>>>>> Stashed changes
```

```{r, fig.width=10}
woody_field_traits_georef %>%
  filter(trait_name %in% core_traits) %>%
  ggplot(aes(lat, value)) + geom_point(alpha = 0.2, stroke=0)+ scale_y_log10() + facet_wrap(~trait_name, scales = "free_y")
```

<<<<<<< Updated upstream
Isaac - looking to do pair-wise relationships for common trait-pairs

```{r}
# pivot trait dataframe to give traits across species from each site
woody_traits_georef_spread <- woody_traits_georef %>% 
  select(observation_id,Temp, Prec, Prec_CV, taxon_name, trait_name, value) %>% 
  pivot_wider(names_from = "trait_name", values_from = "value")%>%
  unnest(trait_names[-12]) %>%
  mutate(Temperature_binary = ifelse(Temp > median(Temp, na.rm=T), paste("Above", round(median(woody_traits_georef$Temp/10, na.rm=T),1), "C",   sep=" ") ,paste("Below", round(median(woody_traits_georef$Temp/10, na.rm=T),1), "C", sep=" "))) %>%
  mutate(Precipitation_binary = ifelse(Prec > median(Prec, na.rm=T), paste("Above", round(median(woody_traits_georef$Prec, na.rm=T),1), "mm",   sep=" ") ,paste("Below", round(median(woody_traits_georef$Prec, na.rm=T),1), "mm", sep=" "))) %>%
  mutate(Precipitation_seasonality_binary = ifelse(Prec_CV > median(Prec_CV, na.rm=T), paste("Above", round(median(woody_traits_georef$Prec_CV, na.rm=T),1),   sep=" ") ,paste("Below", round(median(woody_traits_georef$Prec_CV, na.rm=T),1), sep=" ")))
=======

```{r, fig.width=10}
dev.off()
pdf("output/core_traits_prec.pdf", height=8, width=11)
woody_field_traits_georef %>%
  filter(trait_name %in% core_traits) %>%
  ggplot(aes(Prec, value), col = "black") + geom_point(alpha = 0.2, stroke=0)+scale_x_log10() + scale_y_log10()+facet_wrap(~trait_name, scales = "free_y")
dev.off()
```
>>>>>>> Stashed changes

```{r, fig.width=10}
dev.off()
pdf("output/core_traits_temp.pdf", height=8, width=11)
woody_field_traits_georef %>%
  filter(trait_name %in% core_traits) %>%
  ggplot(aes(Temp, value), col = "black") + geom_point(alpha = 0.2, stroke=0)+scale_x_log10() + scale_y_log10() +facet_wrap(~trait_name, scales = "free_y")
dev.off()
```

```{r, fig.width=10}
dev.off()
pdf("output/core_traits_prec_cv.pdf", height=8, width=11)
woody_field_traits_georef %>%
  filter(trait_name %in% core_traits) %>%
  ggplot(aes(Prec_CV, value), col = "black") + geom_point(alpha = 0.2, stroke=0)+scale_x_log10() + scale_y_log10() +facet_wrap(~trait_name, scales = "free_y")
dev.off()
```

```{r, fig.width=10}
dev.off()
pdf("output/core_traits_DNI.pdf", height=8, width=11)
woody_field_traits_georef %>%
  filter(trait_name %in% core_traits) %>%
  ggplot(aes(DNI, value), col = "black") + geom_point(alpha = 0.2, stroke=0)+scale_x_log10() + scale_y_log10() +facet_wrap(~trait_name, scales = "free_y")
dev.off()
```

<<<<<<< Updated upstream
Plot the 5 most populous pair-wise trait combinations, including high and low climate values - please forgive the coding below, still learning how to do functions using tidyverse and ggplot


```{r echo=FALSE, fig.height= 8, fig.width= 6}
SLA_LA_Temp <- woody_traits_georef_spread %>%
  ggplot(aes(specific_leaf_area, leaf_area, colour=Temperature_binary))+geom_point()+ scale_x_log10() + scale_y_log10()

SLA_LA_Prec <- woody_traits_georef_spread %>%
  ggplot(aes(specific_leaf_area, leaf_area, colour=Precipitation_binary))+geom_point()+ scale_x_log10() + scale_y_log10()

SLA_LA_Prec_CV <- woody_traits_georef_spread %>%
  ggplot(aes(specific_leaf_area, leaf_area, colour=Precipitation_seasonality_binary))+geom_point()+ scale_x_log10() + scale_y_log10()

SLA_LA_Temp / SLA_LA_Prec / SLA_LA_Prec_CV

SLA_LN_Temp <- woody_traits_georef_spread %>%
  ggplot(aes(specific_leaf_area, leaf_N_per_area, colour=Temperature_binary))+geom_point()+ scale_x_log10() + scale_y_log10()

SLA_LN_Prec <- woody_traits_georef_spread %>%
  ggplot(aes(specific_leaf_area, leaf_N_per_area, colour=Precipitation_binary))+geom_point()+ scale_x_log10() + scale_y_log10()

SLA_LN_Prec_CV <- woody_traits_georef_spread %>%
  ggplot(aes(specific_leaf_area, leaf_N_per_area, colour=Precipitation_seasonality_binary))+geom_point()+ scale_x_log10() + scale_y_log10()

SLA_LN_Temp / SLA_LN_Prec / SLA_LN_Prec_CV

SLA_deltaC_Temp <- woody_traits_georef_spread %>%
  ggplot(aes(specific_leaf_area, leaf_delta13C*-1, colour=Temperature_binary))+geom_point()+ scale_x_log10() + scale_y_log10()

SLA_deltaC_Prec <- woody_traits_georef_spread %>%
  ggplot(aes(specific_leaf_area, leaf_delta13C*-1, colour=Precipitation_binary))+geom_point()+ scale_x_log10() + scale_y_log10()

SLA_deltaC_Prec_CV <- woody_traits_georef_spread %>%
  ggplot(aes(specific_leaf_area, leaf_delta13C*-1, colour=Precipitation_seasonality_binary))+geom_point()+ scale_x_log10() + scale_y_log10()

SLA_deltaC_Temp / SLA_deltaC_Prec / SLA_deltaC_Prec_CV

LA_deltaC_Temp <- woody_traits_georef_spread %>%
  ggplot(aes(leaf_delta13C*-1, leaf_area, colour=Temperature_binary))+geom_point()+ scale_x_log10() + scale_y_log10()

LA_deltaC_Prec <- woody_traits_georef_spread %>%
  ggplot(aes(leaf_delta13C*-1, leaf_area, colour=Precipitation_binary))+geom_point()+ scale_x_log10() + scale_y_log10()

LA_deltaC_Prec_CV <- woody_traits_georef_spread %>%
  ggplot(aes(leaf_delta13C*-1, leaf_area, colour=Precipitation_seasonality_binary))+geom_point()+ scale_x_log10() + scale_y_log10()

LA_deltaC_Temp / LA_deltaC_Prec / LA_deltaC_Prec_CV

SLA_WD_Temp <-woody_traits_georef_spread %>%
  ggplot(aes(specific_leaf_area, wood_density, colour=Temperature_binary))+geom_point()+ scale_x_log10() + scale_y_log10()

SLA_WD_Prec <-woody_traits_georef_spread %>%
  ggplot(aes(specific_leaf_area, wood_density, colour=Precipitation_binary))+geom_point()+ scale_x_log10() + scale_y_log10()

SLA_WD_Prec_CV <-woody_traits_georef_spread %>%
  ggplot(aes(specific_leaf_area, wood_density, colour=Precipitation_seasonality_binary))+geom_point()+ scale_x_log10() + scale_y_log10()

SLA_WD_Temp / SLA_WD_Prec / SLA_WD_Prec_CV

```

Compare resolution of Worldclim at highest resolution to CHELSA - has same 19 variables but at higher resolution which could be more useful for ecological applications (see this link https://www.nature.com/articles/sdata2017122) - comparison currently not available as TIF files from CHELSA are local and cannot be committed due to their size. climatedata::get_chelsa allows you to pull the tif files directly from the source but get_chelsa() was returning a connectivity error

```{r echo=FALSE, fig.height= 12, fig.width=12}
# plot_worldclim_climate_data("temp") / plot_chelsa_climate_data("temp")
# 
# plot_worldclim_climate_data("prec") / plot_chelsa_climate_data("prec")

```
=======
```{r, fig.width=10}
dev.off()
pdf("output/core_traits_lat.pdf", height=8, width=11)
woody_field_traits_georef %>%
  filter(trait_name %in% core_traits) %>%
  ggplot(aes(lat, value), col = "black") + geom_point(alpha = 0.2, stroke=0)+ scale_y_log10() +facet_wrap(~trait_name, scales = "free_y")
dev.off()
```


```{r eval=FALSE, fig.width=10, include=FALSE}

woody_field_traits_georef %>%
  filter(trait_name=="leaf_area_calc") %>%
ggplot() +
  aes(x = Temp, y = Prec, z = value) +
  stat_summary_hex(size = 0.1, color = "black", fun = mean) +scale_x_log10() + scale_y_log10() +
  viridis::scale_fill_viridis(trans = "log10") +ggtitle("Leaf area")


woody_field_traits_georef %>%
  filter(trait_name=="leaf_N_per_area_calc") %>%
ggplot() +
  aes(x = Temp, y = Prec, z = value) +
  stat_summary_hex(size = 0.1, color = "black", fun = mean) +scale_x_log10() + scale_y_log10() +
  viridis::scale_fill_viridis(trans = "log10") +ggtitle("Leaf N per area")

woody_field_traits_georef %>%
  filter(trait_name=="plant_height") %>%
ggplot() +
  aes(x = Temp, y = Prec, z = value) +
  stat_summary_hex(size = 0.1, color = "black", fun = mean) +scale_x_log10() + scale_y_log10() +
  viridis::scale_fill_viridis(trans = "log10") +ggtitle("Plant height")

woody_field_traits_georef %>%
  filter(trait_name=="seed_mass") %>%
ggplot() +
  aes(x = Temp, y = Prec, z = value) +
  stat_summary_hex(size = 0.1, color = "black", fun = mean) +scale_x_log10() + scale_y_log10() +
  viridis::scale_fill_viridis(trans = "log10") +ggtitle("Seed mass")

woody_field_traits_georef %>%
  filter(trait_name=="specific_leaf_area_calc") %>%
ggplot() +
  aes(x = Temp, y = Prec, z = value) +
  stat_summary_hex(size = 0.1, color = "black", fun = mean) +scale_x_log10() + scale_y_log10() +
  viridis::scale_fill_viridis(trans = "log10") +ggtitle("Specific leaf area")

woody_field_traits_georef %>%
  filter(trait_name=="wood_density") %>%
ggplot() +
  aes(x = Temp, y = Prec, z = value) +
  stat_summary_hex(size = 0.1, color = "black", fun = mean) +scale_x_log10() + scale_y_log10() +
  viridis::scale_fill_viridis(trans = "log10") +ggtitle("Wood density")
```
>>>>>>> Stashed changes
