---
title: Report on data in AusTraits
output:
  html_document:
    df_print: kable
    highlight: tango
    keep_md: no
    smart: no
    theme: yeti
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options:
  chunk_output_type: console
---

Load data
```{r, warning=FALSE, message=FALSE}
# Install dependencies as needed
# devtools::install_deps()
# Load packages and functions for this analysis

rm(list=ls())
devtools::load_all()
source("R/load_climate_data_AWAP.R")
source("R/compile_and_read_AWAP_data.R")
source("R/load_bioclim.R")
source("R/load_elevation.R")
source("R/sat_slope.R")
source("R/load_worldclim.R")

library(lubridate)
library(R.utils)

# load trait data
austraits <- readRDS("data/austraits-3.0.2.rds")


```

```{r}
sum_no_NA <- function(x) if(all(is.na(x))) NA_integer_ else sum(x, na.rm=TRUE)
```


```{r}
#where are sites located?

location_of_sites <-
  austraits$sites %>%
  filter(site_property %in% c("longitude (deg)", "latitude (deg)")) %>%
  spread(site_property, value) %>%
  rowid_to_column("ID") %>%
  rename(latitude = `latitude (deg)`, longitude = `longitude (deg)`) %>%
  mutate_at(c("latitude", "longitude"), ~suppressWarnings(as.numeric(.x))) %>%
  drop_na(longitude, latitude)


# load map of australia (required for climate data analysis)
au_map <- raster::raster("data/australia.tif")
```

Start with solar which has the highest temporal resolution (i.e. unique daily values)

```{r}
# Finally, extract values for solar irradiance (12, 16-27th November 2009 missing, 2014 incomplete)

#This function compiles extracts point values from daily spatial files and compiles these values into yearly csvs
# compile_daily_solar_data()
#This function reads yearly csvs and compiles them into a total data frame
read_and_combine_compiled_solar_data()

available_years <- c(1990:2008, 2010:2013)

read_csv("data/AWAP_solar/compiled_annual_solar/compiled_total_solar_data.csv") %>%
  filter(year %in% available_years) %>%
  mutate(day = as.numeric(day)) -> solar_data
```

```{r}
#all other data is on at least a monthly scale, so let's convert solar to monthly before joining all enviro variables together (Solar has daily data for 1990 - 2013)

solar_data_monthly <- solar_data %>%
  group_by(ID, year, month) %>%
  summarise(mean_monthly_shortwave_radiation = mean(shortwave_radiation, na.rm=T))

#remove the daily dataset to free up computational resources
rm(solar_data)
```

Next, data for which we have unique monthly data within years.

Start with Australian water availability project data

```{r}
# First, extract values for 9am vapour pressure (hPA) #2020 incomplete

# compile_vp_09_data()

vp_09_data <- read_csv("data/agcd/v1/vapourpres_h09_monthly/mean/r005/01month/compiled_h09_data/compiled_h09_data.csv") %>%
  select(-day) %>%
  filter(year !="2020")

```

```{r}
# Next, extract values for 3pm vapour pressure (hPA) #2020 incomplete

# compile_vp_15_data()

vp_15_data <- read_csv("data/agcd/v1/vapourpres_h15_monthly/compiled_h15_data/compiled_h15_data.csv") %>%
  select(-day)%>%
  filter(year !="2020")

```

```{r}
# compile_annual_rainfall_data()

annual_rainfall_data <- read_csv("data/agcd/v2/r005/01month/compiled_rainfall_data/compiled_rainfall_data.csv") %>%select(-day) %>%
  filter(year !="2020") %>%
  rename(prec_AWAP = prec)
```

```{r}
# compile_max_T_data()

max_t_data <- read_csv("data/agcd/v1/monthly_tmax/compiled_temperature_data/compiled_temperature_data.csv") %>%select(-day) %>%
  filter(year !="2020")
```

Now, data from CRU clim

```{r}
# gunzip("data/CRU/pet/cru_ts4.05.1901.2020.pet.dat.nc.gz", remove=FALSE)
# pet_path <- "data/CRU/pet/cru_ts4.05.1901.2020.pet.dat.nc"
# pet_data <- load_CRU_clim(pet_path)
# write_csv(pet_data, "data/CRU/pet/pet_data.csv")

pet_data <- read_csv("data/CRU/pet/pet_data.csv")  %>%
  rename(pet_CRU = value)%>%
  mutate(date = paste(year, month, day, sep="-")) %>%
  mutate(date = ymd(date)) %>%
  mutate(num_days = days_in_month(date)) %>%
  mutate(monthly_pet_CRU = pet_CRU*num_days) -> pet_data
```


```{r}
# gunzip("data/CRU/temp_avg/cru_ts4.05.1901.2020.tmp.dat.nc.gz", remove=FALSE)
# temp_avg_path <- "data/CRU/temp_avg/cru_ts4.05.1901.2020.tmp.dat.nc"
# temp_avg_data <- load_CRU_clim(temp_avg_path)
# write_csv(temp_avg_data, "data/CRU/temp_avg/temp_avg_data.csv")

temp_avg_data <- read_csv("data/CRU/temp_avg/temp_avg_data.csv") %>%
  rename(temp_avg_CRU = value) %>%
  mutate(sat_slope = sat_slope_func(temp_avg_CRU))
```

```{r}
# gunzip("data/CRU/cloud/cru_ts4.05.1901.2020.cld.dat.nc.gz", remove=FALSE)
# cloud_path <- "data/CRU/cloud/cru_ts4.05.1901.2020.cld.dat.nc"
# cloud_data <- load_CRU_clim(cloud_path)
# write_csv(cloud_data, "data/CRU/cloud/cloud_data.csv")

cloud_data <- read_csv("data/CRU/cloud/cloud_data.csv") %>%
  rename(cloudiness = value)
```


Elevation has just average values for each site, but is needed for SPLASH to calculate net radiation

```{r}
elevation <- load_elevation()
```

```{r}
data_for_splash <- cloud_data %>%
  left_join(temp_avg_data) %>%
  left_join(elevation)

data_for_splash %>%
  mutate(date = paste(year, month, day, sep="-")) %>%
  mutate(date = ymd(date)) %>%
  mutate(doy = yday(date)) %>%
  mutate(sun_frac = (100-cloudiness)/100) -> data_for_splash

write_csv(data_for_splash,"data/splash_input.csv")
```

```{r}
#net radiation from SPLASH

net_rad_paths<-paste("data/net_rad/", list.files("data/net_rad"), sep="")
map(net_rad_paths, read_csv) %>%
  bind_rows() %>%
  select(-sat_slope) %>%
  mutate(net_rad_w = net_rad/24/60/60) -> net_rad 
```


```{r}
left_join(data_for_splash, net_rad) %>%
  left_join(pet_data) %>%
  left_join(annual_rainfall_data) %>% 
  left_join(vp_09_data) %>%
  left_join(vp_15_data) %>%
  left_join(solar_data_monthly) %>%
  left_join(max_t_data) %>%
  mutate(days_in_month = lubridate::days_in_month(date)) %>%
  mutate(eet_day = net_rad_w*sat_slope/(sat_slope+65)/(2.45*1000000)*24*60*60) %>%
  mutate(eet_month = eet_day*days_in_month) -> compiled_monthly_dataset

```


Next, prepare monthly data from the Worldclim dataset. Worldclim has monthly values averaged across years

```{r}
load_worldclim("data/wc2.1_2.5m_tavg/", "avg_temp_wordlclim")

avg_temp_data <- read_csv("data/wc2.1_2.5m_tavg/avg_temp_wordlclim.csv")

avg_temp_data %>%
  group_by(ID) %>%
  summarise(avg_temp_WC = mean(value, na.rm=T)) %>%
  drop_na() -> avg_temp_WC

#growing season temperature excludes months where the average monthly temperature is less than 5

avg_temp_data %>%
  filter(value > 5) %>%
  group_by(ID) %>%
  summarise(avg_gs_temp_WC = mean(value)) -> avg_growing_season_temp_WC

```

```{r}
load_worldclim("data/wc2.1_2.5m_srad/", "solar_wordlclim")

solar_worldclim_data <- read_csv("data/wc2.1_2.5m_srad/solar_wordlclim.csv")

solar_worldclim_data %>%
  group_by(ID) %>%
  summarise(solar_WC = mean(value, na.rm=T)) %>%
  drop_na() -> solar_worldclim
```

```{r}
climate_data <- compiled_monthly_dataset %>%
  left_join(avg_temp_WC) %>%
  left_join(avg_growing_season_temp_WC) %>%
  left_join(solar_worldclim)
```


```{r}
#add vapour pressure deficit to dataframe using formulas for saturated vapour pressure (esat; kPa) and vapour pressure deficit (VPD; kPa) in Monteith JL & Unsworth MH (1990) Principles of environmental physics

es_T_star = 0.611 #kPa
T_dash = 36.0;  #K
T_star = 273.0; #K
A = 17.27
DEG_TO_KELVIN = 273.0 # convert celcius to kelvins
hPa_to_kPA = 1/10 #vapour pressure is in hPa

climate_data$max_t_K<- climate_data$max_t+DEG_TO_KELVIN

climate_data$esat = es_T_star * exp(A * (climate_data$max_t_K - T_star) / (climate_data$max_t_K - T_dash))

#convert vapour pressure to kPa

climate_data$vp_09_kpa <- climate_data$vp09 * hPa_to_kPA
climate_data$vp_15_kpa <- climate_data$vp15 * hPa_to_kPA

#VPD is saturated vapour pressure minus vapour pressure

climate_data$VPD_09 <- climate_data$esat - climate_data$vp_09_kpa
climate_data$VPD_15 <- climate_data$esat - climate_data$vp_15_kpa
```

```{r}
#calculate a yearly value for each site, we want the sum of the monthly sums of rainfall and eet, and the average for the remaining variables

climate_data_yearly_summary <- climate_data %>%
  group_by(year, ID, longitude,latitude) %>%
  summarise(temp_avg_CRU = mean(temp_avg_CRU, na.rm=T),
            temp_avg_WC = mean(avg_temp_WC, na.rm=T),
            temp_avg_gs_WC = mean(avg_gs_temp_WC, na.rm=T), 
            eet_annual = sum_no_NA(eet_month), 
            VPD_09 = mean(VPD_09, na.rm=T), 
            VPD_15 = mean(VPD_15, na.rm=T), 
            annual_prec_AWAP=sum_no_NA(prec_AWAP),
            annual_pet_CRU=sum_no_NA(monthly_pet_CRU),
            mean_annual_shortwave_radiation_AWAP=mean(mean_monthly_shortwave_radiation, na.rm=T), 
            mean_annual_shortwave_radiation_WC = mean(solar_WC,na.rm=T)) %>%
  mutate(MI = annual_prec_AWAP/eet_annual)

```

```{r}
#set up a yearly value for each site
climate_data_site_level_average <- climate_data_yearly_summary %>%
  group_by(ID) %>%
  summarise(VPD_09 = mean(VPD_09, na.rm=T), 
            VPD_15 = mean(VPD_15, na.rm=T), 
            annual_prec_AWAP = mean(annual_prec_AWAP,na.rm=T),
            annual_pet_CRU = mean(annual_pet_CRU, na.rm=T),
            mean_annual_shortwave_radiation_AWAP = mean(mean_annual_shortwave_radiation_AWAP, na.rm=T),
             mean_annual_shortwave_radiation_WC = mean(mean_annual_shortwave_radiation_WC, na.rm=T),
            mean_eet_annual = mean(eet_annual,na.rm=T), 
            mean_MI = mean(MI, na.rm=T), 
            mean_temp_avg_CRU = mean(temp_avg_CRU, na.rm=T),
            mean_temp_avg_WC = mean(temp_avg_WC, na.rm=T),
            mean_temp_avg_gs_WC = mean(temp_avg_gs_WC, na.rm=T))
```


```{r}
#bind wordclim data to compare temp and rainfall data to

#load bioclim data
au_bioclim <- load_climate_data()

au_sites_bioclim <-
  austraits$sites %>%
  filter(site_property %in% c("longitude (deg)", "latitude (deg)")) %>%
  spread(site_property, value) %>%
  rowid_to_column("ID") %>%
  rename(latitude = `latitude (deg)`, longitude = `longitude (deg)`) %>%
  mutate_at(c("latitude", "longitude"), ~suppressWarnings(as.numeric(.x))) %>%
  combine_occurence_climate(au_bioclim)

```

```{r}
combined_climate_data <- left_join(climate_data_site_level_average, au_sites_bioclim)
```

```{r}
combined_climate_data %>%
  ggplot() +
  geom_point(aes(x=VPD_09, y=VPD_15))

combined_climate_data %>%
  ggplot() +
  geom_point(aes(x=Temp_BC, y=mean_temp_avg_CRU))

combined_climate_data %>%
  ggplot() +
  geom_point(aes(x=annual_prec_AWAP, y=Prec_BC))

combined_climate_data %>%
  ggplot() +
  geom_point(aes(x=mean_annual_shortwave_radiation_AWAP, y=mean_annual_shortwave_radiation_WC)) 

combined_climate_data %>%
  ggplot() +
  geom_point(aes(x=annual_prec, y=mean_annual_shortwave_radiation)) + scale_x_log10()

combined_climate_data %>%
  ggplot() +
  geom_point(aes(x=mean_eet_annual, y=annual_pet_CRU))


combined_climate_data %>%
  ggplot() +
  geom_point(aes(x=max_t, y=avg_gs_temp))


```


Calculate leaf_N_per_area and leaf_area for additional observations:
```{r include=FALSE}
#two current problems, in summarise argument, `first` should be `mean` but breaking on instances with a single observation (almost all). Second, this code means the newly calculated values don't have value_type and replicate linked to them
austraits$traits %>% 
  filter(trait_name %in% c("leaf_area","leaf_dry_mass","specific_leaf_area","leaf_N_per_dry_mass",
                           "leaf_N_per_area")) %>%
  select(trait_name, value, observation_id, taxon_name,dataset_id,site_name, context_name, date) %>%
  mutate(value = as.numeric(value)) %>%
  group_by(trait_name, observation_id, taxon_name, dataset_id, site_name, context_name, date) %>%
  summarise(value = mean(value)) %>%
  ungroup() %>%
  spread(key = trait_name, value = value) %>% 
  mutate(specific_leaf_area = as.numeric(specific_leaf_area),
         leaf_N_per_dry_mass = as.numeric(leaf_N_per_dry_mass),
         leaf_area = as.numeric(leaf_area),
         leaf_dry_mass = as.numeric(leaf_dry_mass),
         leaf_N_per_area = as.numeric(leaf_N_per_area)) %>%
  mutate(leaf_N_calc_SLA = leaf_N_per_dry_mass / specific_leaf_area,
         leaf_N_calc_LA = leaf_N_per_dry_mass*leaf_dry_mass/leaf_area,
         leaf_N_per_area_combined = ifelse(!is.na(leaf_N_per_area),leaf_N_per_area,leaf_N_calc_SLA),
         leaf_N_per_area_combined = ifelse(!is.na(leaf_N_per_area_combined),leaf_N_per_area_combined,leaf_N_calc_LA),
         specific_leaf_area_combined = ifelse(!is.na(specific_leaf_area),specific_leaf_area,leaf_area/leaf_dry_mass),
         leaf_area_combined = ifelse(!is.na(leaf_area),leaf_area,specific_leaf_area*leaf_dry_mass)) ->calculated_data

calculated_data %>%
  select(observation_id, taxon_name, dataset_id, site_name, context_name, date, 
         leaf_N_per_area_combined) %>%
  filter(!is.na(leaf_N_per_area_combined)) %>%
  mutate(units = "g/m2",
         trait_name = "leaf_N_per_area_calc",
         value = as.character(leaf_N_per_area_combined)) %>%
  select(-leaf_N_per_area_combined) -> leaf_N_per_area_calcuated

calculated_data %>%
  select(observation_id, taxon_name, dataset_id, site_name, context_name, date, 
         leaf_area_combined) %>%
  filter(!is.na(leaf_area_combined)) %>%
  mutate(units = "mm2",
         trait_name = "leaf_area_calc",
         value = as.character(leaf_area_combined)) %>%
  select(-leaf_area_combined) -> leaf_area_calcuated

calculated_data %>%
  select(observation_id, taxon_name, dataset_id, site_name, context_name, date, 
         specific_leaf_area_combined) %>%
  filter(!is.na(specific_leaf_area_combined)) %>%
  mutate(units = "mm2/mg",
         trait_name = "specific_leaf_area_calc",
         value = as.character(specific_leaf_area_combined)) %>%
  select(-specific_leaf_area_combined) -> specific_leaf_area_calcuated


```

Extract values for ci:ca ratios from delta13c using equations in Farqhuar (1989) and Wang et al. (2017)

Carbon isotope discrimination values are dervied from delta_13C $\delta ^{13}C$ using the following equation from Farqhuar (1989):

$$\Delta = \frac{\delta_a-\delta ^{13}C}{1+\delta ^{13}C/1000}$$

The ci:ca ratio $(X)$ is linked to the carbon isotope discrimination value of plant samples according the following equation:

$$X = \frac{\Delta-a}{b-a},$$
where a’ and b’ have standard values 4.4‰ and 27‰, representing the diffusional 
and biochemical components of carbon isotope discrimination, respectively (Wang et al. 2017).


``` {r}
austraits$traits %>% 
  filter(trait_name %in% c("ci_over_ca", "leaf_delta13C")) %>%
  select(trait_name, value, observation_id, taxon_name,dataset_id,site_name, context_name, date) %>%
  mutate(value = as.numeric(value)) %>%
  group_by(trait_name, observation_id, taxon_name, dataset_id, site_name, context_name, date) %>%
  summarise(value = mean(value)) %>%
  ungroup() %>%
  spread(key = trait_name, value = value)%>%
  mutate(delta_C = (-8-leaf_delta13C)/(1+leaf_delta13C/1000))%>%
  mutate(ci_over_ca_calc = (delta_C-4.4)/(27-4.4)) %>%
  mutate(ci_over_ca_combined = ifelse(!is.na(ci_over_ca),ci_over_ca,ci_over_ca_calc)) ->calculated_data

calculated_data %>%
  select(observation_id, taxon_name, dataset_id, site_name, context_name, date, 
         ci_over_ca_combined) %>%
  filter(!is.na(ci_over_ca_combined)) %>%
  mutate(units = "unitless",
         trait_name = "ci_over_ca_calc",
         value = as.character(ci_over_ca_combined)) %>%
  select(-ci_over_ca_combined) -> ci_over_ca_calculated

```


``` {r, fig.width=10}

calculated_data %>% drop_na(leaf_delta13C, ci_over_ca)%>% mutate(proportion_difference = abs(ci_over_ca - ci_over_ca_calc)/ci_over_ca_calc) -> ci_over_ca_validation


hist(ci_over_ca_validation$proportion_difference, xlab = "abs(Observed_X-Predicted_X)/Predicted_X", main = "Difference between observed and predicted X")
```


``` {r, fig.width=10}
#add a 1:1 line

ci_over_ca_validation %>% ggplot(aes(x=ci_over_ca, y=leaf_delta13C))+geom_point()+theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))

ci_over_ca_validation %>% ggplot(aes(x=ci_over_ca, y=ci_over_ca_calc))+geom_point()+xlab("Observed ci_over_ca")+ylab("Estimated ci_over_ca")+theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))
```


``` {r, fig.width=10}
ci_over_ca_validation %>% ggplot(aes(x=ci_over_ca, y=ci_over_ca_calc, color=dataset_id == "Dong_2017"))+geom_point()+xlab("Observed ci_over_ca")+ylab("Estimated ci_over_ca")+theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))
```

Select subset of traits:
```{r include=FALSE}
traits <- 
  austraits$traits %>%
  bind_rows(leaf_N_per_area_calcuated) %>%
  bind_rows(specific_leaf_area_calcuated) %>%
  bind_rows(ci_over_ca_calculated)%>%
  bind_rows(leaf_area_calcuated) %>%
  mutate(value = as.numeric(value)) %>%
  group_by(dataset_id, taxon_name, site_name, trait_name) %>%
  summarise(value=median(value)) %>%
  ungroup() %>%
  left_join(by = c("dataset_id", "site_name"), combined_climate_data)

```


```{r include=FALSE}
field_traits <- traits %>%
  filter(!(dataset_id %in% c("Ahrens_2019", "Crous_2013", "Crous_2019", "Du_2018",
                            "Du_2019", "Duan_2015", "EsperonRodriguez_2019",
                            "EsperonRodriguez_2020","Everingham_2020",
                            "Farrell_2012", "Farrell_2013", "Farrell_2017", "Firn_2019",
                            "Gallagher_2011_1", "Gallagher_2011_3", "Gallagher_2015", 
                            "Gardiner_2019", "Geange_2017", "Geange_2020", "Ghannoum_2010",
                            "Harrison_2009",
                            "Hassiotou_2009", "Huang_2015", "Jagdish_2020", "Lewis_2015",
                            "Lusk_2012", "Lusk_2014", "Moore_2019_2", "Reynolds_2018",
                            "Roderick_2002", "Schulze_2006", "Smith_2012",
                            "Tomlinson_2013","Tomlinson_2019", "Turner_2010", "Vesk_2019",
                            "Vlasveld_2018", "Warren_2005", "Warren_2006"))) %>%
  filter(!(site_name %in% c("Currency Creek Arboretum","Currency Creek Arboretum, South Australia",
                            "Australian National University glasshouse","MU_campus",
                            "Australian National Botanic Gardens","Australian National University"))) 
```

Create a new trait dataframe to only include observations with associated latitude and longitude
```{r}
field_traits_georef <- 
  field_traits %>%
  drop_na(lat,lon)
```

Define some core traits to focus on

```{r}

core_traits <- c("ci_over_ca_calc","leaf_N_per_area_calc","specific_leaf_area_calc","wood_density","plant_height", "seed_mass", "leaf_area_calc")

field_traits_georef %>%
  filter(trait_name %in% core_traits) -> field_core_traits_georef

```

```{r}

woody_plant_growth_forms <- c("hemi-epiphyte_shrub", "hemi-epiphyte_tree", "leafless_shrub_tree", "parasite_woody", "subshrub", "shrub", "treelet", "tree", "shrub tree", "shrub treelet",
"shrub subshrub","herb subshrub", "herb_large shrub", "herb shrub", "tree tree")

austraits$traits %>% 
  filter(trait_name== "plant_growth_form") %>%
  group_by(taxon_name) %>% 
  filter(all((value %in% woody_plant_growth_forms))) %>%
  ungroup() %>%
  distinct(taxon_name)-> woody_taxa_plant_growth_form

austraits$traits %>% 
  filter(trait_name == "woodiness") %>% 
  group_by(taxon_name) %>%
  filter(all(value %in% c("woody","semi_woody"))) %>%
  distinct(taxon_name)%>%
  bind_rows(woody_taxa_plant_growth_form) %>%
    distinct(taxon_name) -> woody_taxa

# list of species which did not have woody data, updated with data pulled off internet
read_csv("data/no_woody_data_updated.csv") %>%
  filter(woodiness %in% c("woody","semi_woody")) %>%
           distinct(taxon_name) %>%
           bind_rows(woody_taxa) %>%
           distinct(taxon_name) -> woody_taxa

```

Create a new trait dataframe to only include observations with associated latitude and longitude
```{r include=FALSE}
woody_field_core_traits_georef <- 
  field_core_traits_georef %>%
  filter(taxon_name %in% woody_taxa$taxon_name)
```


```{r include=FALSE}
ferns_species <-
  austraits$taxa%>%
  filter(family %in% c("Schizaeaceae", "Gleicheniaceae","Dennstaedtiaceae", "Polypodiaceae","Dicksoniaceae", "Datiscaceae"))

palm_species <-
  austraits$taxa%>%
  filter(family %in% c("Arecaceae"))

cycad_species <-
  austraits$taxa%>%
  filter(family %in% c("Cycadaceae","Zamiaceae", "Stangeriaceae"))

grass_species <- 
  austraits$taxa %>%
  filter(family %in% c("Poaceae"))

ginger_species <- 
  austraits$taxa %>%
  filter(family %in% c("Zingiberaceae"))

woody_field_core_traits_georef_no_ferns_palms_cycads_grass_ginger_schulze1998 <- 
  woody_field_core_traits_georef %>%
  filter(!taxon_name %in% ferns_species$taxon_name) %>%
  filter(!taxon_name %in% palm_species$taxon_name)%>%
  filter(!taxon_name %in% cycad_species$taxon_name)%>%
  filter(!taxon_name %in% grass_species$taxon_name)%>%
  filter(!taxon_name %in% ginger_species$taxon_name) ->
  woody_field_core_traits_georef_no_ferns_palms_cycads_grass_ginger
```

We will now inspect distribution of data in each study relative to total distribution to see if there is any erroneous data
```{r}
trait_by_dataset_core_traits<-function(trait) {woody_field_core_traits_georef_no_ferns_palms_cycads_grass_ginger %>%
  filter(trait_name == trait) %>%
  distinct(dataset_id) -> dataset_index_leaf_area

purrr::map(dataset_index_leaf_area$dataset_id, plot_trait_by_dataset, trait)
}

purrr::map(core_traits, trait_by_dataset_core_traits)
```


```{r}
woody_field_core_traits_georef_no_ferns_palms_cycads_grass_ginger %>%
  group_by(taxon_name, trait_name) %>%
  filter(n()>1)

woody_field_core_traits_georef_no_ferns_palms_cycads_grass_ginger <-
  left_join(woody_field_core_traits_georef_no_ferns_palms_cycads_grass_ginger, summarise_trait_data) %>%
  filter(num_obs > 2) %>%
  mutate(range_lower = mean_value-sd2_value, range_upper = mean_value+sd2_value) %>%
  filter(value < range_lower | value > range_upper)

```


```{r include=FALSE}
%>%
  #remove Schulze_1998 because bad LMA values
  filter(!dataset_id =="Schulze_1998") %>%
  #remove Senecio pinnatifolius, LMA way too high?
  filter(!taxon_name =="Senecio pinnatifolius var. alpinus") %>%
  filter(!(taxon_name == "Atriplex stipitata" & trait_name == "ci_over_ca_calc")) %>%
  filter(!(trait_name == "specific_leaf_area_calc" & value < 0.63))

woody_field_core_traits_georef_no_ferns_palms_cycads_grass_ginger_schulze1998 %>% 
  filter(trait_name == "specific_leaf_area_calc") -> asd

```

```{r}
datasets_with_leaf_area <- read_csv("data/datasets_with_compound_species_master.csv")
# 
datasets_with_leaf_area %>%
  filter(is.na(leaflet_or_leaf)|leaflet_or_leaf == "leaf") %>%
  mutate(compound_species = "leaf") %>%
  select(compound_species, dataset_id) -> leaf_area_studies

woody_field_core_traits_georef_no_ferns_palms_cycads_grass_ginger_schulze1998 %>% 
  left_join(leaf_area_studies) %>% 
  mutate(compound_species = ifelse(is.na(compound_species), "leaflet", compound_species))%>%
  mutate(leaf_was_compound = paste(trait_name, compound_species))%>% 
  filter(leaf_was_compound != "leaf_area_calc leaf")-> woody_field_core_traits_georef_no_ferns_palms_cycads_grass_ginger_schulze19981

```


```{r}
# library(sf)
# library(rgeos)
# library("ggspatial")
# library("rnaturalearth")
# library("rnaturalearthdata")
# 
# world <- ne_countries(scale = "medium", returnclass = "sf")

# gene world map
```


```{r, fig.height==15, fig.width=15}
# ggplot(data = world) +
#   geom_sf() +
#   labs( x = "Longitude", y = "Latitude") +
#   coord_sf(xlim = c(100.00, 160.00), ylim = c(-45.00, -10.00)) +
#   annotation_scale(location = "bl", width_hint = 0.5) +
#   annotation_north_arrow(location = "bl", which_north = "true", 
#                          pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
#                          style = north_arrow_fancy_orienteering) +
#   geom_point(data=woody_field_traits_georef_core_traits, aes(x=lon, y=lat, colour=annual_prec)) +
#   theme_bw() +
  # facet_wrap(~trait_name)
```


```{r, fig.height==15, fig.width=15}
# ggplot(data = world) +
#   geom_sf() +
#   labs( x = "Longitude", y = "Latitude") +
#   coord_sf(xlim = c(100.00, 160.00), ylim = c(-45.00, -10.00)) +
#   annotation_scale(location = "bl", width_hint = 0.5) +
#   annotation_north_arrow(location = "bl", which_north = "true", 
#                          pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
#                          style = north_arrow_fancy_orienteering) +
#   geom_point(data=woody_field_traits_georef_core_traits, aes(x=lon, y=lat, colour=mean_annual_shortwave_radiation)) +
#   theme_bw()+
#   facet_wrap(~trait_name)
```


```{r, fig.height==15, fig.width=15}
# ggplot(data = world) +
#   geom_sf() +
#   labs( x = "Longitude", y = "Latitude") +
#   coord_sf(xlim = c(100.00, 160.00), ylim = c(-45.00, -10.00)) +
#   annotation_scale(location = "bl", width_hint = 0.5) +
#   annotation_north_arrow(location = "bl", which_north = "true", 
#                          pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
#                          style = north_arrow_fancy_orienteering) +
#   geom_point(data=woody_field_traits_georef_core_traits, aes(x=lon, y=lat, colour=max_t)) +
#   theme_bw()+
#   facet_wrap(~trait_name)
```


```{r, fig.height==15, fig.width=15}
# ggplot(data = world) +
#   geom_sf() +
#   labs( x = "Longitude", y = "Latitude") +
#   coord_sf(xlim = c(100.00, 160.00), ylim = c(-45.00, -10.00)) +
#   annotation_scale(location = "bl", width_hint = 0.5) +
#   annotation_north_arrow(location = "bl", which_north = "true", 
#                          pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
#                          style = north_arrow_fancy_orienteering) +
#   geom_point(data=woody_field_traits_georef_core_traits, aes(x=lon, y=lat, colour=VPD_15)) +
#   theme_bw()+
#   facet_wrap(~trait_name)

```


Basic patterns with environment

```{r}

core_traits <- c("ci_over_ca_calc","leaf_N_per_area_calc","specific_leaf_area_calc","wood_density","plant_height","leaf_area_calc")

```

```{r, fig.width=10}
library("wesanderson")
pal<-wesanderson::wes_palette("Darjeeling1", 5)

trait_names <- tibble(trait_name = core_traits,
                      trait = c("ci:ca","leaf_N", "SLA", "Wood_dens", "height", "leaf_area"))

woody_field_core_traits_georef_no_ferns_palms_cycads_grass_ginger_schulze19981 %>%
  filter(trait_name %in% core_traits) %>%
  left_join(by="trait_name", trait_names) %>%
  select(trait, value, Temp = mean_temp_avg_CRU, VPD = VPD_15, PAR = mean_annual_shortwave_radiation_AWAP, MI = mean_MI) %>%
  pivot_longer(-c("trait", "value"), names_to = "env", values_to = "env_value") %>%
  mutate(env = factor(env, levels = c("MI", "Temp", "VPD","PAR"))) -> data_for_plot


png("output/trait_by_environment_viridis_high_res_small_axes.png", height=1200, width=2000, res=300)
data_for_plot %>% 
  ggplot(aes(env_value, value, col = trait)) + geom_point(alpha = 0.3, shape=16, stroke=0, show.legend = F) + scale_y_log10()   +scale_x_log10()+ theme_classic() + facet_grid(trait ~ env, scales = "free")+ scale_colour_viridis(discrete = TRUE) +theme(axis.title.x=element_blank()) +theme(axis.title.y=element_blank()) +
  theme(strip.text.x = element_text(size = 6), strip.text.y = element_text(size = 6), axis.text=element_text(size=7))
dev.off()


# woody_field_core_traits_georef_no_ferns_palms_cycads_grass_ginger_schulze1998 %>%
#   filter(trait_name %in% "leaf_area_calc") %>%
#   mutate(moisture_index=cut(moisture_index, breaks=c(-Inf, 0.5, 0.8, Inf), labels=c("Dry","Intermediate","Wet"))) %>%
#   # left_join(by="trait_name", trait_names) %>%
#   select(trait_name, value, Lat = lat, MI = moisture_index) %>%
#   # pivot_longer(-c("trait", "value"), names_to = "env", values_to = "env_value") %>%
#   # mutate(env = factor(env, levels = c("Lat", "MI"))) %>%
#   ggplot(aes(x=Lat, y=value)) + geom_point() + scale_y_log10() + facet_wrap(~MI)
```


```{r, fig.width=10}

woody_field_core_traits_georef_no_ferns_palms_cycads_grass_ginger_schulze1998 %>%
  filter(trait_name %in% core_traits) %>%
  left_join(by="trait_name", trait_names) %>%
  filter(grepl("Eucalyptus|Corymbia|Angophora", taxon_name)) %>%
  select(trait, value, Temp = mean_temp_avg, VPD = VPD_15, PAR = mean_annual_shortwave_radiation, MI = mean_MI) %>%
  pivot_longer(-c("trait", "value"), names_to = "env", values_to = "env_value") %>%
  mutate(env = factor(env, levels = c("MI", "Temp", "VPD","PAR"))) -> euc_data_for_plot

png("output/trait_by_environment_euc_viridis_high_res_small_axes.png", height=1200, width=2000, res=300)
data_for_plot %>% 
  ggplot(aes(env_value, value, col = trait)) + geom_point(alpha = 0.30, shape=16, stroke=0, show.legend = F, color = "gray") + geom_point(data = euc_data_for_plot, aes(env_value, value, col = trait),alpha = 0.30, shape=19, stroke=0, show.legend = F) + scale_colour_viridis(discrete = TRUE)+ scale_y_log10()   +scale_x_log10()+ theme_classic() + facet_grid(trait ~ env, scales = "free") +theme(axis.title.x=element_blank()) +theme(axis.title.y=element_blank()) +
  theme(strip.text.x = element_text(size = 6), strip.text.y = element_text(size = 6), axis.text=element_text(size=7)) 
dev.off()

```